### 2025-02-14 paper/main.tex
init

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2024

% ready for submission
\usepackage{neurips_2024}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
%     \usepackage[preprint]{neurips_2024}

% to compile a camera-ready version, add the [final] option, e.g.:
%     \usepackage[final]{neurips_2024}

% to avoid loading the natbib package, add option nonatbib:
%    \usepackage[nonatbib]{neurips_2024}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors

\title{Unveiling the Syntax Within: Interpreting Grammar Embeddings in Meta’s LLaMA Models}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

\author{%
  Pratim Chowdhary\thanks{Use footnote for providing further information
    about author (webpage, alternative address)---\emph{not} for acknowledging
    funding agencies.} \\
  Department of Computer Science\\
  Dartmouth College\\
  \texttt{cpratim.25@dartmouth.edu} \\
  % examples of more authors
  \And
  Peter Chin \\
  Department of Engineering\\
  Thayer School of Engineering\\
  \texttt{pc@dartmouth.edu} \\
  % \AND
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
}

\begin{document}

\maketitle

\begin{abstract}
  The abstract paragraph should be indented \nicefrac{1}{2}~inch (3~picas) on
  both the left- and right-hand margins. Use 10~point type, with a vertical
  spacing (leading) of 11~points.  The word \textbf{Abstract} must be centered,
  bold, and in point size 12. Two line spaces precede the abstract. The abstract
  must be limited to one paragraph.
\end{abstract}

\section{Submission of papers to the M3L Workshop at NeurIPS 2024}

Please read the instructions below carefully and follow them faithfully.

\subsection{Style}

Papers to be submitted to the Mathematics of Modern Machine Learning (M3L)
Workshop at NeurIPS 2024 must be prepared according to the instructions
presented here.

Authors are required to use the Mathematics of Modern Machine Learning (M3L) \LaTeX{} style files obtainable at the
workshop website \url{https://sites.google.com/view/m3l-2024/call-for-papers}. Please make sure you use the current files
and not previous versions. Tweaking the style files may be grounds for
rejection.

\subsection{Retrieval of style files}

The style files for the Mathematics of Modern Machine Learning (M3L) Workshop at NeurIPS 2024 and other conference information are available on
the website at
\begin{center}
  \url{https://sites.google.com/view/m3l-2024}
\end{center}
The file \verb+main.pdf+ contains these instructions and illustrates the
various formatting requirements your NeurIPS paper must satisfy.

The \LaTeX{} style file contains three optional arguments: \verb+final+, which
creates a camera-ready copy, \verb+preprint+, which creates a preprint for
submission to, e.g., arXiv, and \verb+nonatbib+, which will not load the
\verb+natbib+ package for you in case of package clash.

\paragraph{Preprint option}
If you wish to post a preprint of your work online, e.g., on arXiv, using the
NeurIPS style, please use the \verb+preprint+ option. This will create a
nonanonymized version of your work with the text ``Preprint. Work in progress.''
in the footer. This version may be distributed as you see fit, as long as you do not say which conference it was submitted to. Please \textbf{do
  not} use the \verb+final+ option, which should \textbf{only} be used for
papers accepted to the Mathematics of Modern Machine Learning (M3L) Workshop at NeurIPS 2024.

At submission time, please omit the \verb+final+ and \verb+preprint+
options. This will anonymize your submission and add line numbers to aid
review. Please do \emph{not} refer to these line numbers in your paper as they
will be removed during generation of camera-ready copies.

The file \verb+main.tex+ may be used as a ``shell'' for writing your
paper. All you have to do is replace the author, title, abstract, and text of
the paper with your own.

The formatting instructions contained in these style files are summarized in
Sections \ref{gen_inst}, \ref{headings}, and \ref{others} below.

\section{General formatting instructions}
\label{gen_inst}

The text must be confined within a rectangle 5.5~inches (33~picas) wide and
9~inches (54~picas) long. The left margin is 1.5~inch (9~picas).  Use 10~point
type with a vertical spacing (leading) of 11~points.  Times New Roman is the
preferred typeface throughout, and will be selected for you by default.
Paragraphs are separated by \nicefrac{1}{2}~line space (5.5 points), with no
indentation.

The paper title should be 17~point, initial caps/lower case, bold, centered
between two horizontal rules. The top rule should be 4~points thick and the
bottom rule should be 1~point thick. Allow \nicefrac{1}{4}~inch space above and
below the title to rules. All pages should start at 1~inch (6~picas) from the
top of the page.

For the final version, authors' names are set in boldface, and each name is
centered above the corresponding address. The lead author's name is to be listed
first (left-most), and the co-authors' names (if different address) are set to
follow. If there is only one co-author, list both author and co-author side by
side.

Please pay special attention to the instructions in Section \ref{others}
regarding figures, tables, acknowledgments, and references.

\section{Headings: first level}
\label{headings}

All headings should be lower case (except for first word and proper nouns),
flush left, and bold.

First-level headings should be in 12-point type.

\subsection{Headings: second level}

Second-level headings should be in 10-point type.

\subsubsection{Headings: third level}

Third-level headings should be in 10-point type.

\paragraph{Paragraphs}

There is also a \verb+\paragraph+ command available, which sets the heading in
bold, flush left, and inline with the text, with the heading followed by 1\,em
of space.

\section{Citations, figures, tables, references}
\label{others}

These instructions apply to everyone.

\subsection{Citations within the text}

The \verb+natbib+ package will be loaded for you by default.  Citations may be
author/year or numeric, as long as you maintain internal consistency.  As to the
format of the references themselves, any style is acceptable as long as it is
used consistently.

The documentation for \verb+natbib+ may be found at
\begin{center}
  \url{http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}
\end{center}
Of note is the command \verb+\citet+, which produces citations appropriate for
use in inline text.  For example,
\begin{verbatim}
   \citet{hasselmo} investigated\dots
\end{verbatim}
produces
\begin{quote}
  Hasselmo, et al.\ (1995) investigated\dots
\end{quote}

If you wish to load the \verb+natbib+ package with options, you may add the
following before loading the \verb+neurips_2024+ package:
\begin{verbatim}
   \PassOptionsToPackage{options}{natbib}
\end{verbatim}

If \verb+natbib+ clashes with another package you load, you can add the optional
argument \verb+nonatbib+ when loading the style file:
\begin{verbatim}
   \usepackage[nonatbib]{neurips_2024}
\end{verbatim}

As submission is double blind, refer to your own published work in the third
person. That is, use ``In the previous work of Jones et al.\ [4],'' not ``In our
previous work [4].'' If you cite your other papers that are not widely available
(e.g., a journal paper under review), use anonymous author names in the
citation, e.g., an author of the form ``A.\ Anonymous'' and include a copy of the anonymized paper in the supplementary material.

\subsection{Footnotes}

Footnotes should be used sparingly.  If you do require a footnote, indicate
footnotes with a number\footnote{Sample of the first footnote.} in the
text. Place the footnotes at the bottom of the page on which they appear.
Precede the footnote with a horizontal rule of 2~inches (12~picas).

Note that footnotes are properly typeset \emph{after} punctuation
marks.\footnote{As in this example.}

\subsection{Figures}

\begin{figure}
  \centering
  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
\end{figure}

All artwork must be neat, clean, and legible. Lines should be dark enough for
purposes of reproduction. The figure number and caption always appear after the
figure. Place one line space before the figure caption and one line space after
the figure. The figure caption should be lower case (except for first word and
proper nouns); figures are numbered consecutively.

You may use color figures.  However, it is best for the figure captions and the
paper body to be legible if the paper is printed in either black/white or in
color.

\subsection{Tables}

All tables must be centered, neat, clean and legible.  The table number and
title always appear before the table.  See Table~\ref{sample-table}.

Place one line space before the table title, one line space after the
table title, and one line space after the table. The table title must
be lower case (except for first word and proper nouns); tables are
numbered consecutively.

Note that publication-quality tables \emph{do not contain vertical rules.} We
strongly suggest the use of the \verb+booktabs+ package, which allows for
typesetting high-quality, professional tables:
\begin{center}
  \url{https://www.ctan.org/pkg/booktabs}
\end{center}
This package was used to typeset Table~\ref{sample-table}.

\begin{table}
  \caption{Sample table title}
  \label{sample-table}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{2}{c}{Part}                   \\
    \cmidrule(r){1-2}
    Name     & Description     & Size ($\mu$m) \\
    \midrule
    Dendrite & Input terminal  & $\sim$100     \\
    Axon     & Output terminal & $\sim$10      \\
    Soma     & Cell body       & up to $10^6$  \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Math}
Note that display math in bare TeX commands will not create correct line numbers for submission. Please use LaTeX (or AMSTeX) commands for unnumbered display math. (You really shouldn't be using \$\$ anyway; see \url{https://tex.stackexchange.com/questions/503/why-is-preferable-to} and \url{https://tex.stackexchange.com/questions/40492/what-are-the-differences-between-align-equation-and-displaymath} for more information.)

\subsection{Final instructions}

Do not change any aspects of the formatting parameters in the style files.  In
particular, do not modify the width or length of the rectangle the text should
fit into, and do not change font sizes (except perhaps in the
\textbf{References} section; see below). Please note that pages should be
numbered.

\section{Preparing PDF files}

Please prepare submission files with paper size ``US Letter,'' and not, for
example, ``A4.''

Fonts were the main cause of problems in the past years. Your PDF file must only
contain Type 1 or Embedded TrueType fonts. Here are a few instructions to
achieve this.

\begin{itemize}

\item You should directly generate PDF files using \verb+pdflatex+.

\item You can check which fonts a PDF files uses.  In Acrobat Reader, select the
  menu Files$>$Document Properties$>$Fonts and select Show All Fonts. You can
  also use the program \verb+pdffonts+ which comes with \verb+xpdf+ and is
  available out-of-the-box on most Linux machines.

\item \verb+xfig+ "patterned" shapes are implemented with bitmap fonts.  Use
  "solid" shapes instead.

\item The \verb+\bbold+ package almost always uses bitmap fonts.  You should use
  the equivalent AMS Fonts:
\begin{verbatim}
   \usepackage{amsfonts}
\end{verbatim}
followed by, e.g., \verb+\mathbb{R}+, \verb+\mathbb{N}+, or \verb+\mathbb{C}+
for $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.  You can also use the following
workaround for reals, natural and complex:
\begin{verbatim}
   \newcommand{\RR}{I\!\!R} %real numbers
   \newcommand{\Nat}{I\!\!N} %natural numbers
   \newcommand{\CC}{I\!\!\!\!C} %complex numbers
\end{verbatim}
Note that \verb+amsfonts+ is automatically loaded by the \verb+amssymb+ package.

\end{itemize}

If your file contains type 3 fonts or non embedded TrueType fonts, we will ask
you to fix it.

\subsection{Margins in \LaTeX{}}

Most of the margin problems come from figures positioned by hand using
\verb+\special+ or other commands. We suggest using the command
\verb+\includegraphics+ from the \verb+graphicx+ package. Always specify the
figure width as a multiple of the line width as in the example below:
\begin{verbatim}
   \usepackage[pdftex]{graphicx} ...
   \includegraphics[width=0.8\linewidth]{myfile.pdf}
\end{verbatim}
See Section 4.4 in the graphics bundle documentation
(\url{http://mirrors.ctan.org/macros/latex/required/graphics/grfguide.pdf})

A number of width problems arise when \LaTeX{} cannot properly hyphenate a
line. Please give LaTeX hyphenation hints using the \verb+\-+ command when
necessary.

\begin{ack}
Use unnumbered first level headings for the acknowledgments. All acknowledgments
go at the end of the paper before the list of references. Moreover, you are required to declare
funding (financial activities supporting the submitted work) and competing interests (related financial activities outside the submitted work).
More information about this disclosure can be found at: \url{https://neurips.cc/Conferences/2024/PaperInformation/FundingDisclosure}.

Do {\bf not} include this section in the anonymized submission, only in the final paper. You can use the \texttt{ack} environment provided in the style file to autmoatically hide this section in the anonymized submission.
\end{ack}

\section*{References}

References follow the acknowledgments in the camera-ready paper. Use unnumbered first-level heading for
the references. Any choice of citation style is acceptable as long as you are
consistent. It is permissible to reduce the font size to \verb+small+ (9 point)
when listing the references.
Note that the Reference section does not count towards the page limit.
\medskip

{
\small

[1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms for
connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and T.K.\ Leen
(eds.), {\it Advances in Neural Information Processing Systems 7},
pp.\ 609--616. Cambridge, MA: MIT Press.

[2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS: Exploring
  Realistic Neural Models with the GEneral NEural SImulation System.}  New York:
TELOS/Springer--Verlag.

[3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of learning and
recall at excitatory recurrent synapses and cholinergic modulation in rat
hippocampal region CA3. {\it Journal of Neuroscience} {\bf 15}(7):5249-5262.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix

\section{Appendix / supplemental material}

Optionally include supplemental material (complete proofs, additional experiments and plots) in appendix.

\end{document}

### 2025-02-14 helpers.py

@@ -1,0 +1,79 @@
+import torch
+from transformers import AutoTokenizer, AutoModelForCausalLM, pipeline
+import pandas as pd
+import json
+import os
+from tqdm import tqdm
+
+device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
+
+def load_model(model_id="meta-llama/Llama-3.2-3B"):
+
+    tokenizer = AutoTokenizer.from_pretrained(model_id)
+    model = AutoModelForCausalLM.from_pretrained(
+        model_id,
+        torch_dtype=torch.float16, 
+        device_map=device
+    )
+
+    model.to(device)
+    tokenizer.pad_token = tokenizer.eos_token
+    return model, tokenizer
+
+
+def load_dataset_from_file(grammar_file, model, tokenizer, batch_size=100):
+
+    with open(grammar_file, 'r') as f:
+        data = [json.loads(line) for line in f]
+
+    positive = [l['sentence_good'] for l in data]
+    negative = [l['sentence_bad'] for l in data]
+
+    positive_encodings, negative_encodings = [], []
+
+    for i in tqdm(range(0, len(positive), batch_size)):
+
+        with torch.no_grad():
+            positive_tokens = tokenizer(positive[i:i+batch_size], return_tensors='pt', padding=True, truncation=True)
+            positive_tokens = {k: v.to(device) for k, v in positive_tokens.items()}
+
+            positive_outputs = model(
+                **positive_tokens,
+                max_length=1,
+                output_hidden_states=True,
+                return_dict=True
+            )
+            positive_hidden_states = positive_outputs.hidden_states[-1][:, -1, :]
+
+            negative_tokens = tokenizer(negative[i:i+batch_size], return_tensors='pt', padding=True, truncation=True)
+            negative_tokens = {k: v.to(device) for k, v in negative_tokens.items()}
+            negative_outputs = model(
+                **negative_tokens,
+                max_length=1,
+                output_hidden_states=True,
+                return_dict=True
+            )
+            negative_hidden_states = negative_outputs.hidden_states[-1][:, -1, :]
+            
+            positive_encodings.append(positive_hidden_states)
+            negative_encodings.append(negative_hidden_states)
+
+    positive_encodings = torch.cat(positive_encodings, dim=0)
+    negative_encodings = torch.cat(negative_encodings, dim=0)
+    return positive_encodings, negative_encodings
+
+
[…]

### 2025-02-14 util/helpers.py
restructure

@@ -15,7 +15,6 @@ def load_model(model_id="meta-llama/Llama-3.2-3B"):
         torch_dtype=torch.float16, 
         device_map=device
     )
-
     model.to(device)
     tokenizer.pad_token = tokenizer.eos_token
     return model, tokenizer
@@ -63,7 +62,7 @@ def load_dataset_from_file(grammar_file, model, tokenizer, batch_size=100):
     return positive_encodings, negative_encodings
 
 
-def load_complete_dataset(data_dir='blimp/data/', batch_size=100, n_files=None):
+def load_complete_dataset(data_dir='../data/blimp/', batch_size=100, n_files=None):
     model, tokenizer = load_model()
     
     positive_encodings, negative_encodings = [], []

### 2025-02-14 util/model.py

@@ -1,0 +1,16 @@
+from transformers import AutoTokenizer, AutoModelForCausalLM, pipeline
+import torch
+
+device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
+
+def load_model(model_id="meta-llama/Llama-3.2-3B"):
+
+    tokenizer = AutoTokenizer.from_pretrained(model_id)
+    model = AutoModelForCausalLM.from_pretrained(
+        model_id,
+        torch_dtype=torch.float16, 
+        device_map=device
+    )
+    model.to(device)
+    tokenizer.pad_token = tokenizer.eos_token
+    return model, tokenizer

### 2025-02-23 paper/main.tex
abstract

@@
   \texttt{pc@dartmouth.edu} \\
+  \And
+  Deepernab Chakrabarty \\
+  Department of Computer Science\\
+  Dartmouth College\\
+  \texttt{deepernab@dartmouth.edu} \\
   % \AND
@@
 \begin{abstract}
-  The abstract paragraph should be indented \nicefrac{1}{2}~inch (3~picas) on
-  both the left- and right-hand margins. Use 10~point type, with a vertical
-  spacing (leading) of 11~points.  The word \textbf{Abstract} must be centered,
-  bold, and in point size 12. Two line spaces precede the abstract. The abstract
-  must be limited to one paragraph.
+  This paper investigates the mechanisms by which large language models (LLMs) encode 
+  grammatical knowledge, focusing on Meta's LLaMA models. By leveraging embedding vectors, 
+  we classify grammatically correct sentences and analyze the activation patterns of 
+  attention heads to identify their roles in processing specific grammatical structures. 
+  Furthermore, we explore the effects of selectively removing these attention heads, shedding 
+  light on how grammar is embedded within the model's architecture. Our findings aim to enhance
+  the understanding of LLMs' linguistic capabilities and their internal organization of syntactic knowledge.
 \end{abstract}
@@
 
-Please read the instructions below carefully and follow them faithfully.
+Please read the instructions below carefully and follow them faithfully. fds
 

### 2025-02-27 paper/main.tex
graphs

@@
 
-\section{Submission of papers to the M3L Workshop at NeurIPS 2024}
+\section{Introduction}
 
-Please read the instructions below carefully and follow them faithfully. fds
+\section{Related Work}
 
-\subsection{Style}
+\begin{enumerate}
 
-Papers to be submitted to the Mathematics of Modern Machine Learning (M3L)
-Workshop at NeurIPS 2024 must be prepared according to the instructions
-presented here.
+  \item \textbf{Deciphering Stereotypes in Pre-Trained Language Models} 
+  \begin{itemize}
+    \item This paper goes into details about how stereotypes are encoded in the embeddings of pre-trained language models and how specific attention heads are responsible for encoding them.
+  \end{itemize}
+  \item \textbf{}
 
-Authors are required to use the Mathematics of Modern Machine Learning (M3L) \LaTeX{} style files obtainable at the
-workshop website \url{https://sites.google.com/view/m3l-2024/call-for-papers}. Please make sure you use the current files
-and not previous versions. Tweaking the style files may be grounds for
-rejection.
+\end{enumerate}
 
-\subsection{Retrieval of style files}
+\section{Methodology}
 
-The style files for the Mathematics of Modern Machine Learning (M3L) Workshop at NeurIPS 2024 and other conference information are available on
-the website at
-\begin{center}
-  \url{https://sites.google.com/view/m3l-2024}
-\end{center}
-The file \verb+main.pdf+ contains these instructions and illustrates the
-various formatting requirements your NeurIPS paper must satisfy.
+\section{Experiments}
 
-The \LaTeX{} style file contains three optional arguments: \verb+final+, which
-creates a camera-ready copy, \verb+preprint+, which creates a preprint for
-submission to, e.g., arXiv, and \verb+nonatbib+, which will not load the
-\verb+natbib+ package for you in case of package clash.
+\section{Discussion}
 
-\paragraph{Preprint option}
-If you wish to post a preprint of your work online, e.g., on arXiv, using the
-NeurIPS style, please use the \verb+preprint+ option. This will create a
-nonanonymized version of your work with the text ``Preprint. Work in progress.''
-in the footer. This version may be distributed as you see fit, as long as you do not say which conference it was submitted to. Please \textbf{do
-  not} use the \verb+final+ option, which should \textbf{only} be used for
-papers accepted to the Mathematics of Modern Machine Learning (M3L) Workshop at NeurIPS 2024.
-
-At submission time, please omit the \verb+final+ and \verb+preprint+
-options. This will anonymize your submission and add line numbers to aid
-review. Please do \emph{not} refer to these line numbers in your paper as they
-will be removed during generation of camera-ready copies.
-
-The file \verb+main.tex+ may be used as a ``shell'' for writing your
-paper. All you have to do is replace the author, title, abstract, and text of
-the paper with your own.
-
-The formatting instructions contained in these style files are summarized in
-Sections \ref{gen_inst}, \ref{headings}, and \ref{others} below.
-
-\section{General formatting instructions}
-\label{gen_inst}
-
-The text must be confined within a rectangle 5.5~inches (33~picas) wide and
-9~inches (54~picas) long. The left margin is 1.5~inch (9~picas).  Use 10~point
-type with a vertical spacing (leading) of 11~points.  Times New Roman is the
-preferred typeface throughout, and will be selected for you by default.
-Paragraphs are separated by \nicefrac{1}{2}~line space (5.5 points), with no
-indentation.
-
-The paper title should be 17~point, initial caps/lower case, bold, centered
-between two horizontal rules. The top rule should be 4~points thick and the
-bottom rule should be 1~point thick. Allow \nicefrac{1}{4}~inch space above and
-below the title to rules. All pages should start at 1~inch (6~picas) from the
-top of the page.
-
-For the final version, authors' names are set in boldface, and each name is
-centered above the corresponding address. The lead author's name is to be listed
-first (left-most), and the co-authors' names (if different address) are set to
-follow. If there is only one co-author, list both author and co-author side by
-side.
-
-Please pay special attention to the instructions in Section \ref{others}
-regarding figures, tables, acknowledgments, and references.
-
-\section{Headings: first level}
-\label{headings}
-
-All headings should be lower case (except for first word and proper nouns),
-flush left, and bold.
-
-First-level headings should be in 12-point type.
-
-\subsection{Headings: second level}
-
-Second-level headings should be in 10-point type.
-
-\subsubsection{Headings: third level}
-
-Third-level headings should be in 10-point type.
-
-\paragraph{Paragraphs}
-
-There is also a \verb+\paragraph+ command available, which sets the heading in
-bold, flush left, and inline with the text, with the heading followed by 1\,em
-of space.
-
-\section{Citations, figures, tables, references}
-\label{others}
-
-These instructions apply to everyone.
-
-\subsection{Citations within the text}
-
-The \verb+natbib+ package will be loaded for you by default.  Citations may be
-author/year or numeric, as long as you maintain internal consistency.  As to the
-format of the references themselves, any style is acceptable as long as it is
-used consistently.
-
-The documentation for \verb+natbib+ may be found at
-\begin{center}
-  \url{http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}
-\end{center}
-Of note is the command \verb+\citet+, which produces citations appropriate for
-use in inline text.  For example,
-\begin{verbatim}
-   \citet{hasselmo} investigated\dots
-\end{verbatim}
-produces
-\begin{quote}
-  Hasselmo, et al.\ (1995) investigated\dots
-\end{quote}
-
-If you wish to load the \verb+natbib+ package with options, you may add the
-following before loading the \verb+neurips_2024+ package:
-\begin{verbatim}
-   \PassOptionsToPackage{options}{natbib}
-\end{verbatim}
-
-If \verb+natbib+ clashes with another package you load, you can add the optional
-argument \verb+nonatbib+ when loading the style file:
-\begin{verbatim}
-   \usepackage[nonatbib]{neurips_2024}
-\end{verbatim}
-
-As submission is double blind, refer to your own published work in the third
-person. That is, use ``In the previous work of Jones et al.\ [4],'' not ``In our
-previous work [4].'' If you cite your other papers that are not widely available
-(e.g., a journal paper under review), use anonymous author names in the
-citation, e.g., an author of the form ``A.\ Anonymous'' and include a copy of the anonymized paper in the supplementary material.
-
-\subsection{Footnotes}
-
-Footnotes should be used sparingly.  If you do require a footnote, indicate
-footnotes with a number\footnote{Sample of the first footnote.} in the
-text. Place the footnotes at the bottom of the page on which they appear.
-Precede the footnote with a horizontal rule of 2~inches (12~picas).
-
-Note that footnotes are properly typeset \emph{after} punctuation
-marks.\footnote{As in this example.}
-
-\subsection{Figures}
-
-\begin{figure}
-  \centering
-  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
-  \caption{Sample figure caption.}
-\end{figure}
-
-All artwork must be neat, clean, and legible. Lines should be dark enough for
-purposes of reproduction. The figure number and caption always appear after the
-figure. Place one line space before the figure caption and one line space after
-the figure. The figure caption should be lower case (except for first word and
-proper nouns); figures are numbered consecutively.
-
-You may use color figures.  However, it is best for the figure captions and the
-paper body to be legible if the paper is printed in either black/white or in
-color.
-
-\subsection{Tables}
-
-All tables must be centered, neat, clean and legible.  The table number and
-title always appear before the table.  See Table~\ref{sample-table}.
-
-Place one line space before the table title, one line space after the
-table title, and one line space after the table. The table title must
-be lower case (except for first word and proper nouns); tables are
-numbered consecutively.
-
-Note that publication-quality tables \emph{do not contain vertical rules.} We
-strongly suggest the use of the \verb+booktabs+ package, which allows for
-typesetting high-quality, professional tables:
-\begin{center}
-  \url{https://www.ctan.org/pkg/booktabs}
-\end{center}
-This package was used to typeset Table~\ref{sample-table}.
-
-\begin{table}
-  \caption{Sample table title}
-  \label{sample-table}
-  \centering
-  \begin{tabular}{lll}
-    \toprule
-    \multicolumn{2}{c}{Part}                   \\
-    \cmidrule(r){1-2}
-    Name     & Description     & Size ($\mu$m) \\
-    \midrule
-    Dendrite & Input terminal  & $\sim$100     \\
-    Axon     & Output terminal & $\sim$10      \\
-    Soma     & Cell body       & up to $10^6$  \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-\subsection{Math}
-Note that display math in bare TeX commands will not create correct line numbers for submission.
-Please use LaTeX (or AMSTeX) commands for unnumbered display math.
-(You really shouldn't be using \$\$ anyway; see \url{https://tex.stackexchange.com/questions/503/why-is-preferable-to} and \url{https://tex.stackexchange.com/questions/40492/what-are-the-differences-between-align-equation-and-displaymath} for more information.)
-
-\subsection{Final instructions}
-
-Do not change any aspects of the formatting parameters in the style files.  In
-particular, do not modify the width or length of the rectangle the text should
-fit into, and do not change font sizes (except perhaps in the
-\textbf{References} section; see below). Please note that pages should be
-numbered.
-
-\section{Preparing PDF files}
-
-Please prepare submission files with paper size ``US Letter,'' and not, for
-example, ``A4.''
-
-Fonts were the main cause of problems in the past years. Your PDF file must only
-contain Type 1 or Embedded TrueType fonts. Here are a few instructions to
-achieve this.
-
-\begin{itemize}
-
-\item You should directly generate PDF files using \verb+pdflatex+.
-
-\item You can check which fonts a PDF files uses.  In Acrobat Reader, select the
-  menu Files$>$Document Properties$>$Fonts and select Show All Fonts. You can
-  also use the program \verb+pdffonts+ which comes with \verb+xpdf+ and is
-  available out-of-the-box on most Linux machines.
-
-\item \verb+xfig+ "patterned" shapes are implemented with bitmap fonts.  Use
-  "solid" shapes instead.
-
-\item The \verb+\bbold+ package almost always uses bitmap fonts.  You should use
-  the equivalent AMS Fonts:
-\begin{verbatim}
-   \usepackage{amsfonts}
-\end{verbatim}
-followed by, e.g., \verb+\mathbb{R}+, \verb+\mathbb{N}+, or \verb+\mathbb{C}+
-for $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.  You can also use the following
-workaround for reals, natural and complex:
-\begin{verbatim}
-   \newcommand{\RR}{I\!\!R} %real numbers
-   \newcommand{\Nat}{I\!\!N} %natural numbers
-   \newcommand{\CC}{I\!\!\!\!C} %complex numbers
-\end{verbatim}
-Note that \verb+amsfonts+ is automatically loaded by the \verb+amssymb+ package.
-
-\end{itemize}
-
-If your file contains type 3 fonts or non embedded TrueType fonts, we will ask
-you to fix it.
-
-\subsection{Margins in \LaTeX{}}
-
-Most of the margin problems come from figures positioned by hand using
-\verb+\special+ or other commands. We suggest using the command
-\verb+\includegraphics+ from the \verb+graphicx+ package. Always specify the
-figure width as a multiple of the line width as in the example below:
-\begin{verbatim}
-   \usepackage[pdftex]{graphicx} ...
-   \includegraphics[width=0.8\linewidth]{myfile.pdf}
-\end{verbatim}
-See Section 4.4 in the graphics bundle documentation
-(\url{http://mirrors.ctan.org/macros/latex/required/graphics/grfguide.pdf})
-
-A number of width problems arise when \LaTeX{} cannot properly hyphenate a
-line. Please give LaTeX hyphenation hints using the \verb+\-+ command when
-necessary.
-
-\begin{ack}
-Use unnumbered first level headings for the acknowledgments. All acknowledgments
-go at the end of the paper before the list of references. Moreover, you are required to declare
-funding (financial activities supporting the submitted work) and competing interests (related financial activities outside the submitted work).
-More information about this disclosure can be found at: \url{https://neurips.cc/Conferences/2024/PaperInformation/FundingDisclosure}.
-
-Do {\bf not} include this section in the anonymized submission, only in the final paper.
-You can use the \texttt{ack} environment provided in the style file to autmoatically hide this section in the anonymized submission.
-\end{ack}
+\section{Conclusion}
 

### 2025-02-27 src/load.py

@@ -1,0 +1,105 @@
+from util.data import load_multiple_choice_dataset, load_binary_dataset, load_raw_embeddings_dataset
+from transformers import AutoTokenizer, AutoModelForCausalLM
+import torch
+import pickle
+import os
+import json
+
+device = torch.device("cuda:7" if torch.cuda.is_available() else "cpu")
+
+def get_device_memory_report(device):
+    print(f'Device: {device} [{torch.cuda.get_device_name(device)}]')
+    free_memory, total_memory = torch.cuda.mem_get_info(device)
+    
+    free_memory_gb = free_memory / (1024 ** 3)
+    total_memory_gb = total_memory / (1024 ** 3)
+    
+    print(f"Free Memory: {free_memory_gb:.2f}/{total_memory_gb:.2f} GB [{free_memory / total_memory * 100:.2f}%]")
+
+
+if __name__ == "__main__":
+    get_device_memory_report(device)
+
+    output_dir = '../data/blimp-output'
+
+    files = [
+        '../data/blimp/anaphor_number_agreement.jsonl'
+    ]
+
+    model_ids = [
+        'meta-llama/Llama-3.2-1B-Instruct',
+        'meta-llama/Llama-3.2-3B-Instruct',
+        'meta-llama/Llama-3.1-8B-Instruct'
+    ]
+
+    n_shots = [
+        1, 5, 10, 20
+    ]
+
+    for model_id in model_ids:
+        model_name = model_id.split('/')[-1]
+
+        model = AutoModelForCausalLM.from_pretrained(
+            model_id,
+            device_map=device,
+            torch_dtype=torch.float16,
+        )
+        model.to(device)
+        tokenizer = AutoTokenizer.from_pretrained(
+            model_id,
+            device_map=device,
+            torch_dtype=torch.float16,
+        )
+        tokenizer.pad_token = tokenizer.eos_token 
+        for file in files:
+            file_name = file.split('/')[-1].split('.')[0]
+
+            raw_embeddings = load_raw_embeddings_dataset(file, model, tokenizer, device, limit=None)  
+            with open(f'{output_dir}/{model_name}/{file_name}_raw_embeddings.pkl', 'wb') as f:
+                pickle.dump(raw_embeddings, f)
+
+            print(f'Saved {file_name}_raw_embeddings.pkl')
+            print('-' * 100)
+
+            torch.cuda.empty_cache()
+        
+            for n_shot in n_shots:
[…]

### 2025-02-27 src/util/data.py

@@ -1,0 +1,285 @@
+import torch
+import torch.nn as nn
+import torch.optim as optim
+import torch.nn.functional as F
+from warnings import filterwarnings
+
+import json
+import pandas as pd
+import numpy as np
+import logging
+
+from tqdm import tqdm
+from random import shuffle, uniform
+from transformers import pipeline
+
+logging.getLogger("transformers").setLevel(logging.ERROR)
+filterwarnings('ignore')
+
+__system_message_multiple_choice = """
+You are a helpful assistant that will help me understand which sentence is grammatically correct. 
+You should only output the number of the correct sentence: 1 / 2
+"""
+
+__system_message_binary = """
+You are a helpful assistant that will help me understand if a sentence is grammatically correct. 
+You should only output '1' if the sentence is grammatically correct, and '0' if it is not.
+"""
+
+__question_format_multiple_choice = """
+Which sentence is more grammatical and native-like? 
+1) {sentence_1} 
+2) {sentence_2}
+"""
+
+__question_format_binary = """
+Is this sentence grammatically correct? 
+1) {sentence_1}
+"""
+
+
+def load_grammar_data(file_path):
+    with open(file_path, 'r') as f:
+        grammar_data = [json.loads(line) for line in f]
+    return grammar_data
+
+
+def load_LLM_messages_multiple_choice(grammar_data, n_shot=5):
+
+    messages = [
+        {
+            "role": "system",
+            "content": __system_message_multiple_choice
+        }
+    ]
+    shuffle(grammar_data)
+    for idx in range(n_shot):
+        good_sentence = grammar_data[idx]['sentence_good']
+        bad_sentence = grammar_data[idx]['sentence_bad']
+        messages.extend([
+            {
+                "role": "user",
+                "content": __question_format_multiple_choice.format(
+                    sentence_1=good_sentence, 
+                    sentence_2=bad_sentence
+                )
+            },
+            {
+                "role": "assistant",
+                "content": '1'
+            },
+            {
+                "role": "user",
+                "content": __question_format_multiple_choice.format(
+                    sentence_1=bad_sentence, 
+                    sentence_2=good_sentence
+                )
+            },
+            {
+                "role": "assistant",
+                "content": '2'
+            }
+        ])
+    return messages
+
+
+def load_LLM_messages_binary(grammar_data, n_shot=5):
+    messages = [
+        {
+            "role": "system",
+            "content": __system_message_binary
+        }
[…]

### 2025-02-27 src/util/helpers.py

@@ -5,28 +5,61 @@ import json
 import os
 from tqdm import tqdm
 
-device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
+device = torch.device("cuda:1" if torch.cuda.is_available() else "cpu")
 
-def load_model(model_id="meta-llama/Llama-3.2-3B"):
+
+def load_blimp_(grammar_file):
+    with open(grammar_file, 'r') as f:
+        data = [json.loads(line) for line in f]
+
+    positive = [l['sentence_good'] for l in data]
+    negative = [l['sentence_bad'] for l in data]
+
+    return positive, negative
+
+
+def get_blimp_dataset(data_dir='../data/blimp/'):
+    
+    positive, negative = [], []
+    for file in os.listdir(data_dir):
+        if file.endswith('.jsonl'):
+            with open(os.path.join(data_dir, file), 'r') as f:
+                data = [json.loads(line) for line in f]
+
+            positive.extend([l['sentence_good'] for l in data])
+            negative.extend([l['sentence_bad'] for l in data])
+    return positive, negative
+
+
+def load_model(model_id="meta-llama/Llama-3.2-1B", device="cuda:1"):
 
     tokenizer = AutoTokenizer.from_pretrained(model_id)
     model = AutoModelForCausalLM.from_pretrained(
         model_id,
         torch_dtype=torch.float16, 
-        device_map=device
+        device_map=device,
+        attn_implementation="eager"
     )
     model.to(device)
     tokenizer.pad_token = tokenizer.eos_token
     return model, tokenizer
 
 
-def load_dataset_from_file(grammar_file, model, tokenizer, batch_size=100):
+def load_dataset_from_file(grammar_file, model, tokenizer, add_grammar_prompt=False, batch_size=100):
 
     with open(grammar_file, 'r') as f:
         data = [json.loads(line) for line in f]
 
     positive = [l['sentence_good'] for l in data]
     negative = [l['sentence_bad'] for l in data]
+
+    if add_grammar_prompt:
+        positive = [
+            "Output 'YES' if the following sentence is grammatically correct, otherwise output 'NO': " + p for p in positive
+        ]
+        negative = [
+            "Output 'YES' if the following sentence is grammatically correct, otherwise output 'NO': " + n for n in negative
+        ]
 
     positive_encodings, negative_encodings = [], []
 
@@ -59,16 +92,17 @@ def load_dataset_from_file(grammar_file, model, tokenizer, batch_size=100):
 
[…]

### 2025-04-22 paper/main.tex
new

@@
 \usepackage{xcolor}         % colors
+\usepackage{graphicx}       % for images
+\usepackage{adjustbox}      % for figure positioning
+\usepackage{amsmath}        % for align environment
 
-\title{Unveiling the Syntax Within: Interpreting Grammar Embeddings in Meta’s LLaMA Models}
-
-% The \author macro works with any number of authors. There are two commands
-% used to separate the names and addresses of multiple authors: \And and \AND.
-%
-% Using \And between authors leaves it to LaTeX to determine where to break the
-% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
-% authors names on the first line, and the last on the second line, try using
-% \AND instead of \And before the third author name.
+\title{Unveiling the Syntax Within: Interpreting Grammar Embeddings in Small-Scale LLMs}
 
@@
   \texttt{deepernab@dartmouth.edu} \\
-  % \AND
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-  % \And
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-  % \And
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
 }
@@
 \begin{abstract}
-  This paper investigates the mechanisms by which large language models (LLMs) encode 
-  grammatical knowledge, focusing on Meta's LLaMA models. By leveraging embedding vectors, 
-  we classify grammatically correct sentences and analyze the activation patterns of 
-  attention heads to identify their roles in processing specific grammatical structures. 
-  Furthermore, we explore the effects of selectively removing these attention heads, shedding 
-  light on how grammar is embedded within the model's architecture. Our findings aim to enhance
-  the understanding of LLMs' linguistic capabilities and their internal organization of syntactic knowledge.
+  We investigate how grammatical knowledge is embedded within small language models (1-8B parameters), an understudied yet practically important class of LLMs.
+Through systematic analysis of Meta's Llama, Gemma, and Qwen model families using the BLiMP benchmark of linguistic minimal pairs, we examine grammatical representations across different architectures and parameter scales.
+Our findings reveal that grammatical knowledge scales non-uniformly, with certain model families demonstrating specific grammatical strengths independent of size.
+We observe complex grammatical phenomena showing non-linear scaling patterns, while others plateau quickly.
+Notably, smaller models from certain architectures sometimes outperform larger models from different families on specific grammatical tasks, suggesting architectural inductive biases significantly influence grammatical knowledge acquisition.
+This work provides insights into grammar representation in contemporary LLMs with implications for model design and enhancement of grammatical competence in resource-constrained deployment scenarios.
   \end{abstract}
@@
 \section{Introduction}
+Large language models (LLMs) have transformed natural language processing, demonstrating unprecedented capabilities across a spectrum of linguistic tasks—from simple question answering to complex reasoning, creative writing, and code generation \citep{brown2020language, chowdhery2022palm, touvron2023llama2, jiang2023mistral}.
+These models, trained through self-supervised learning on vast corpora of text data, have increasingly approached human-like performance in generating coherent, contextually appropriate, and grammatically well-formed text, despite having no explicit grammatical rules programmed into their architecture.
+This emergent grammatical competence—arising purely from statistical patterns in training data—represents a fascinating case study in how linguistic knowledge can be acquired implicitly through exposure rather than explicit instruction.
 
-\section{Related Work}
+The field has witnessed exponential growth in model size, from early transformer models with hundreds of millions of parameters \citep{vaswani2017attention} to modern giants like GPT-4 \citep{openai2023gpt4} that likely contain trillions of parameters.
+While these massive models have captured headlines with their impressive capabilities, a parallel revolution has been unfolding in the development of smaller, more efficient models in the 1-8 billion parameter range.
+Models like Llama 2 \citep{touvron2023llama2}, Gemma 7B \citep{jiang2023mistral}, and Qwen \citep{bai2023qwen} have demonstrated remarkable performance despite their relatively modest size, making them particularly valuable for practical applications where computational efficiency and deployment costs are significant concerns.
 
-\begin{enumerate}
+These smaller LLMs offer a compelling balance between capability and efficiency, enabling deployment on consumer hardware, edge devices, and resource-constrained environments.
+Their reduced inference costs make them attractive for commercial applications, while their smaller memory footprint allows for fine-tuning and adaptation with more modest computational resources.
+Understanding how these models encode and represent grammatical knowledge is therefore not merely an academic exercise but has significant practical implications for developing more capable, efficient, and linguistically robust language technologies.
 
-  \item \textbf{Deciphering Stereotypes in Pre-Trained Language Models} 
-  \begin{itemize}
-    \item This paper goes into details about how stereotypes are encoded in the embeddings of pre-trained language models and how specific attention heads are responsible for encoding them.
-  \end{itemize}
-  \item \textbf{}
+\section{Related Work and Background}
 
-\end{enumerate}
+\subsection{Grammatical Evaluation of Language Models}
+Research on grammatical capabilities in neural language models has evolved from early work on LSTM-based models \citep{linzen2016assessing} to transformer-based architectures \citep{goldberg2019assessing, devlin2019bert}.
+Evaluation frameworks have progressed from simple subject-verb agreement tests to comprehensive benchmarks like CoLA \citep{warstadt2019neural} and BLiMP \citep{warstadt2020blimp}.
+The BLiMP benchmark, comprising 67 minimal pair tests across 12 linguistic phenomena, has been particularly valuable for assessing grammatical competence.
+Recent studies have extended evaluation to larger models, revealing that while performance scales with size, challenges remain with complex hierarchical structures and garden-path sentences \citep{thrush2022winograd, qian2022limitations}.
+
+\subsection{Scale and Grammatical Competence}
+The relationship between model scale and grammatical abilities follows complex patterns beyond the general power-law scaling observed in language modeling \citep{kaplan2020scaling}.
+Research suggests that rare grammatical constructions require disproportionately more training data \citep{wei2021frequency}, and certain phenomena show nonlinear improvements at specific parameter thresholds \citep{zhang2023language}.
+Studies on compositional generalization indicate that scaling alone may not capture human-like grammatical productivity without architectural innovations \citep{hu2020systematic}.
+Beyond raw parameter count, specific architectural components significantly impact grammatical knowledge acquisition, with attention mechanisms crucial for long-distance dependencies and feed-forward networks encoding categorical information \citep{patel2022mapping}.
+
+\subsection{Probing Language Models for Linguistic Knowledge}
+Researchers have developed various probing techniques to understand how linguistic knowledge is represented within model parameters.
+Structural probes have revealed that models implicitly encode hierarchical syntactic structures \citep{hewitt2019structural}, with syntactic information appearing in earlier layers and semantic information in later ones \citep{tenney2019bert}.
+Transformer-based models exhibit emergent symbolic manipulation capabilities resembling discrete linguistic rules \citep{manning2020emergent}, with individual neurons specializing in specific linguistic features \citep{geva2021transformer}.
+Attention patterns often correspond to syntactic dependencies \citep{levy2018emergence}, with different attention heads specializing in distinct linguistic phenomena \citep{clark2019does}.
+
+\subsection{Small LLMs and Comparative Studies}
+The proliferation of open-weight models like OPT \citep{zhang2022opt}, Llama \citep{touvron2023llama}, Gemma \citep{jiang2023mistral}, and Qwen \citep{bai2023qwen} has enabled more systematic analyses of how capabilities scale and how architectural choices affect performance.
+Studies show that smaller models with innovative architectures can outperform larger ones \citep{jiang2023mistral}, and data quality may be as important as scale for robust grammatical representations \citep{bai2023qwen}.
+Comparative analyses across architectures remain limited but suggest significant variations in performance even at similar parameter scales \citep{talmor2020olmpics, zhao2023survey}.
+Recent frameworks for quantifying evaluation uncertainties \citep{roberts2023quantifying} and structured evaluation approaches \citep{xia2023structured} have revealed that architectural design choices impact specific grammatical phenomena differently, suggesting that model architectures encode grammatical knowledge in fundamentally different ways.
 
 \section{Methodology}
+\subsection{Evaluating Grammatical Knowledge Across Linguistic Categories}
+
+Our evaluation framework divides grammatical phenomena into five key linguistic categories: island effects, quantifiers, ellipsis, argument structure, and binding constraints.
+This categorization allows us to assess whether grammatical knowledge is acquired uniformly or if certain phenomena are represented differently across model architectures and scales.
+
+To visualize how these grammatical categories are embedded within model representations, we applied t-SNE 
+dimensionality reduction to final-layer activations from ungrammatical sentences. 
+Figure~\ref{fig:tsne} shows these representations for three 1B-scale models. The visualizations
+ reveal distinct organizational patterns, with varying degrees of cluster separation suggesting that 
+ models encode grammatical phenomena differently. Llama 1B forms more distinct category clusters, while 
+ Qwen 1.5B shows more integrated representations across categories. 
+\begin{figure}[t]
+  \hspace*{-1cm}
+  \includegraphics[width=1.1\textwidth]{figures/grammar_tsne}
+  \caption{T-SNE visualizations of grammatical representations across three 1B-scale language models.
+Points represent ungrammatical sentences colored by linguistic category.
+Contours indicate density of representations, revealing how different architectures organize grammatical knowledge.}
+  \label{fig:tsne}
+\end{figure}
+
+\subsection{Experimental Protocol for Grammatical Assessment}
+
+We evaluate grammatical knowledge in LLMs through a systematically designed protocol that captures both explicit and implicit grammatical understanding.
+Our methodology encompasses two complementary evaluation paradigms:
+
+\begin{itemize}
+  \item \textbf{Binary Classification}: Models evaluate the grammaticality of individual sentences through explicit prompted judgment.
+This tests the model's ability to directly assess syntactic well-formedness.
+
+  \item \textbf{Minimal Pair Discrimination}: Models select between grammatical and ungrammatical alternatives that differ minimally in structure.
+This approach aligns with psycholinguistic methodologies for investigating human grammatical intuitions.
+\end{itemize}
+
+The prompt templates for each evaluation paradigm are formalized as:
+
+\begin{align}
+  \mathcal{P}_{\text{binary}} &= \text{``Is this sentence grammatically correct/incorrect?: } [S]\text{''} \\
+  \mathcal{P}_{\text{pair}} &= \text{``Which sentence is grammatically better? (A) } [S_1] \text{ (B) } [S_2]\text{''}
+\end{align}
+
+Our analysis extends beyond surface-level responses to probe the internal representations formed during grammatical processing.
+For each sentence evaluation, we extract activation vectors from all $n$ transformer layers, enabling a fine-grained examination of how grammatical knowledge emerges and propagates through the network hierarchy:
+
+\begin{align}
+  \mathbf{A}_s &= \{\mathbf{a}_{s,l_1}, \mathbf{a}_{s,l_2}, \ldots, \mathbf{a}_{s,l_n}\}
+\end{align}
+
+where $\mathbf{A}_s$ represents the complete set of layer-wise activations for sentence $s$. These activation patterns serve as a neural signature of the model's grammatical processing.
+
+To quantify and compare grammatical sensitivity across different architectural layers, we compute the Euclidean distance between activation patterns elicited by grammatical ($g$) and ungrammatical ($u$) sentence pairs:
+
+\begin{align}
+  \text{GramDist}(l_i) &= \frac{1}{|P|}\sum_{(g,u) \in P} \|\mathbf{a}_{g,l_i} - \mathbf{a}_{u,l_i}\|_2
+\end{align}
+
+where $P$ denotes the set of all grammatical/ungrammatical sentence pairs in our evaluation corpus.
+This metric captures the degree to which each layer's representations distinguish between well-formed and ill-formed syntactic structures.
+
+Figure~\ref{fig:layer_sensitivity} illustrates the layer-wise grammatical sensitivity across different model families, revealing architectural variations in grammatical knowledge acquisition.
+
+% \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=0.95\textwidth]{figures/layer_sensitivity}
+%     \caption{Layer-wise grammatical sensitivity across Llama, Gemma, and Qwen model families.
+Higher values indicate stronger differentiation between grammatical and ungrammatical sentences.
+Gemma models demonstrate earlier grammatical differentiation in middle layers, while Llama models show more pronounced differentiation in later layers.}
+%     \label{fig:layer_sensitivity}
+% \end{figure}
 
@@
 
-\section*{References}
-
-References follow the acknowledgments in the camera-ready paper. Use unnumbered first-level heading for
-the references. Any choice of citation style is acceptable as long as you are
-consistent. It is permissible to reduce the font size to \verb+small+ (9 point)
-when listing the references.
-Note that the Reference section does not count towards the page limit.
-\medskip
-
-{
-\small
-
-[1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms for
-connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and T.K.\ Leen
-(eds.), {\it Advances in Neural Information Processing Systems 7},
-pp.\ 609--616. Cambridge, MA: MIT Press.
-
-[2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS: Exploring
-  Realistic Neural Models with the GEneral NEural SImulation System.}  New York:
-TELOS/Springer--Verlag.
-
-[3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of learning and
-recall at excitatory recurrent synapses and cholinergic modulation in rat
-hippocampal region CA3. {\it Journal of Neuroscience} {\bf 15}(7):5249-5262.
-}
+\bibliographystyle{plainnat}
+\bibliography{references}
 
@@
 \appendix
-
 \section{Appendix / supplemental material}

### 2025-04-22 src/module/blocked_attention.py

@@ -4,20 +4,115 @@ from transformers.models.llama.modeling_llama import (
     apply_rotary_pos_emb, 
     eager_attention_forward, 
     logger, 
-    ALL_ATTENTION_FUNCTIONS
+    ALL_ATTENTION_FUNCTIONS,
+    repeat_kv
 )
-
+import numpy as np
 from typing import Tuple, Optional, Unpack, Callable, Dict, List
 import torch
+import torch.nn as nn
+from util.model import clear_cuda_cache
 
 from transformers.cache_utils import Cache
 from transformers.modeling_flash_attention_utils import FlashAttentionKwargs
 
 
-class CustomAttention(LlamaAttention):
-    def __init__(self, config, layer_idx, disable_idx=[]):
+def print_matrix(matrix: np.ndarray):
+    for row in matrix:
+        for col in row:
+            print(f"{col:.2f}", end=" ")
+        print()
+
+
+def increase_entropy(distribution, power=2, percentile=0.2):
+    mean = torch.mean(distribution)
+    deviation = distribution - mean
+    
+    over_mean = torch.clamp(torch.sign(deviation) * (deviation) ** power, 0, 1)
+    under_mean = torch.clamp(torch.sign(deviation) * (deviation) ** power, -1, 0)
+    
+    over_sum = torch.sum(over_mean)
+    under_sum = torch.sum(under_mean)
+    
+    sub = torch.zeros_like(distribution)
+    add = torch.zeros_like(distribution)
+    
+    if over_sum > 0:
+        sub = (over_mean / over_sum) * percentile
+    
+    if under_sum < 0:
+        add = (under_mean / under_sum) * percentile
+    
+    new_dist = distribution - sub + add
+    return new_dist
+
+
+def increase_entropy_batch(attn_weights, power=2, percentile=0.1):
+    # Extract the last row for all samples and heads
+    last_row = attn_weights[:, :, -1, :]
+    # Calculate deviation from mean (along the token dimension)
+    mean = torch.mean(last_row, dim=-1, keepdim=True)
+    deviation = last_row - mean
+    # Calculate components above and below mean
+    over_mean = torch.clamp(torch.sign(deviation) * (deviation) ** power, 0, 1)
+    under_mean = torch.clamp(torch.sign(deviation) * (deviation) ** power, -1, 0)
+    # Handle potential division by zero with safe division
+    over_sum = torch.sum(over_mean, dim=-1, keepdim=True)
[…]

### 2025-04-22 src/scripts/load_grammar.py

@@ -1,92 +1,0 @@
-from util.grammar import load_multiple_choice_dataset, load_binary_dataset, load_raw_embeddings_dataset
-from util.model import load_model, get_device_memory_report
-import torch
-import pickle
-import os
-import json
-
-device = torch.device("cuda:7" if torch.cuda.is_available() else "cpu")
-
-if __name__ == "__main__":
-    get_device_memory_report(device)
-
-    output_dir = '../data/blimp-output'
-
-    files = [
-        '../data/blimp/complex_NP_island.jsonl',
-        # '../data/blimp/determiner_noun_agreement_1.jsonl',
-        # '../data/blimp/determiner_noun_agreement_2.jsonl',
-        # '../data/blimp/adjunct_island.jsonl',
-        # '../data/blimp/anaphor_number_agreement.jsonl'
-    ]
-
-    model_ids = [
-        'meta-llama/Llama-3.2-1B-Instruct',
-        'meta-llama/Llama-3.2-3B-Instruct',
-        'meta-llama/Llama-3.1-8B-Instruct'
-    ]
-
-    n_shots = [
-        0, 1, 5, 10
-    ]
-
-    LIMIT = None
-
-    for model_id in model_ids:
-        model_name = model_id.split('/')[-1]
-
-        model, tokenizer = load_model(model_id, device)
-
-        if not os.path.exists(f'{output_dir}/{model_name}'):
-            os.makedirs(f'{output_dir}/{model_name}')
-
-        tokenizer.pad_token = tokenizer.eos_token 
-        for file in files:
-            grammar_name = file.split('/')[-1].split('.')[0]
-
-            if not os.path.exists(f'{output_dir}/{model_name}/{grammar_name}'):
-                os.makedirs(f'{output_dir}/{model_name}/{grammar_name}')
-
-            raw_embeddings = load_raw_embeddings_dataset(file, model, tokenizer, device, limit=LIMIT)  
-            with open(f'{output_dir}/{model_name}/{grammar_name}/no_prompt_embeddings.pkl', 'wb') as f:
-                pickle.dump(raw_embeddings, f)
-
-            print(f'Saved {grammar_name}/no_prompt_embeddings.pkl')
-            print('-' * 100)
-
-            torch.cuda.empty_cache()
-        
-            for n_shot in n_shots:
-                multiple_choice_dataset, multiple_choice_model_outputs = load_multiple_choice_dataset(file, model, tokenizer, device, n_shot, limit=LIMIT)
[…]

### 2025-04-22 src/scripts/load_hidden.py

@@ -1,51 +1,0 @@
-from util.hidden import load_hidden_states
-from util.model import load_model, get_device_memory_report, clear_cuda_cache
-import torch
-import os
-import pickle
-
-
-device = torch.device("cuda:7" if torch.cuda.is_available() else "cpu")
-
-if __name__ == "__main__":
-    get_device_memory_report(device)
-
-    output_dir = '../data/blimp-hidden'
-
-    files = [
-        '../data/blimp/complex_NP_island.jsonl',
-        '../data/blimp/determiner_noun_agreement_1.jsonl',
-        '../data/blimp/determiner_noun_agreement_2.jsonl',
-        '../data/blimp/adjunct_island.jsonl',
-        '../data/blimp/anaphor_number_agreement.jsonl'
-    ]
-
-    model_ids = [
-        # 'meta-llama/Llama-3.2-1B-Instruct',
-        # 'meta-llama/Llama-3.2-3B-Instruct',
-        # 'mistralai/Mistral-7B-Instruct-v0.3',
-        # 'meta-llama/Llama-3.1-8B-Instruct',
-        'Qwen/Qwen2.5-7B-Instruct-1M'
-    ]
-
-    for model_id in model_ids:
-        print(f'Loading model [{model_id}]')
-        print('-' * 100)
-        model_name = model_id.split('/')[-1]
-        model, tokenizer = load_model(model_id, device, eager_load=True)
-
-        if not os.path.exists(f'{output_dir}/{model_name}'):
-            os.makedirs(f'{output_dir}/{model_name}')
-
-        for file in files:
-            print(f'Loading hidden states for [{file}]')
-            grammar_name = file.split('/')[-1].split('.')[0]
-            
-            hidden_states = load_hidden_states(file, model, tokenizer, device)
-            with open(f'{output_dir}/{model_name}/{grammar_name}.pkl', 'wb') as f:
-                pickle.dump(hidden_states, f)
-
-            print(f'Saved {grammar_name}.pkl')
-            print('-' * 100)
-
-            clear_cuda_cache()

### 2025-04-22 src/util/chat.py

@@ -1,0 +1,270 @@
+from warnings import filterwarnings
+import logging
+from util.data import load_grammar_data
+
+from tqdm import tqdm
+from random import shuffle
+from transformers import pipeline
+import torch
+import numpy as np
+from pprint import pprint
+
+logging.getLogger("transformers").setLevel(logging.ERROR)
+filterwarnings("ignore")
+
+np.random.seed(42)
+
+__system_message_multiple_choice_grammar = """You are a helpful assistant that will help me figure out which sentence is grammatically correct and native-like.
+You should only output the number of the correct sentence, '1' if the first sentence is correct, '2' if the second sentence is correct.
+
+ONLY OUTPUT THE NUMBER OF THE CORRECT SENTENCE, NOTHING ELSE.
+
+For example:
+Which sentence is more grammatical and native-like?
+1) The cat is on the mat.
+2) The mat on is the cat.
+
+2"""
+
+__system_message_binary_grammar = """You are a helpful assistant that will help me figure out if a sentence is grammatically correct and native-like. 
+You should only output '1' if the sentence is grammatically correct and native-like, and '0' if it is not.
+
+ONLY OUTPUT '1' OR '0', NOTHING ELSE.
+
+For example:
+Is this sentence grammatically correct?
+1) The cat is on the mat.
+
+1"""
+
+__question_format_multiple_choice_grammar = """Which sentence is more grammatical and native-like? 
+1) {sentence_1} 
+2) {sentence_2}
+
+ONLY OUTPUT THE NUMBER OF THE CORRECT SENTENCE (1 OR 2), NOTHING ELSE."""
+
+__question_format_binary_grammar = """Is this sentence grammatically correct? 
+1) {sentence_1}
+
+ONLY OUTPUT '1' OR '0', NOTHING ELSE."""
+
+__system_message_multiple_choice_arithmetic = """You are a helpful assistant that will help me determine which of two arithmetic expressions is correct.
+
+ONLY OUTPUT THE NUMBER OF THE CORRECT EXPRESSION (1 OR 2), NOTHING ELSE.
+For example:
+Which of the following arithmetic expressions is correct?
+1) 2 + 2 = 4
+2) 2 + 2 = 5
+
+2"""
+
+__question_format_multiple_choice_arithmetic = """Which of the following arithmetic expressions is correct?
+1) {expression_1}
+2) {expression_2}
+
[…]

### 2025-04-22 src/util/data.py

@@ -1,10 +1,37 @@
 import json
 import pickle
+import os
+import numpy as np
+from tqdm import tqdm
+
 
 def load_grammar_data(file_path):
     with open(file_path, 'r') as f:
         grammar_data = [json.loads(line) for line in f]
     return grammar_data
+
+
+def load_combined_grammar_data(data_path: str = '../data/blimp', n_sample: int = 100):
+    combined_grammar_data = []
+    for file in tqdm(os.listdir(data_path)):
+        if file.endswith('.jsonl'):
+            grammar_data = load_grammar_data(os.path.join(data_path, file))
+            if n_sample is not None:
+                combined_grammar_data.extend(np.random.choice(grammar_data, size=n_sample, replace=False))
+            else:
+                combined_grammar_data.extend(grammar_data)
+    return combined_grammar_data
+
+
+def load_grammar_data_by_category(data_path: str = '../data/blimp', category: str = 'linguistics_term', value: str = 'island_effects', sample_rate: float = 1.0, use_tqdm: bool = False):
+    combined_grammar_data = []
+    for file in tqdm(os.listdir(data_path)) if use_tqdm else os.listdir(data_path):
+        if file.endswith('.jsonl'):
+            grammar_data = load_grammar_data(os.path.join(data_path, file))
+            for example in grammar_data:
+                if example[category] == value and np.random.rand() < sample_rate:
+                    combined_grammar_data.append(example)
+    return combined_grammar_data
 
 
 def load_hidden_data(model_name, grammar_name, data_path='../data/blimp-hidden'):

### 2025-04-22 src/util/grammar.py

@@ -1,263 +1,0 @@
-from warnings import filterwarnings
-import logging
-from util.data import load_grammar_data
-
-from tqdm import tqdm
-from random import shuffle, uniform
-from transformers import pipeline
-import json
-import torch
-
-logging.getLogger("transformers").setLevel(logging.ERROR)
-filterwarnings('ignore')
-
-__system_message_multiple_choice = """
-You are a helpful assistant that will help me figure out which sentence is grammatically correct and native-like.
-You should only output the number of the correct sentence, '1' if the first sentence is correct, '2' if the second sentence is correct.
-"""
-
-__system_message_binary = """
-You are a helpful assistant that will help me figure out if a sentence is grammatically correct and native-like. 
-You should only output '1' if the sentence is grammatically correct and native-like, and '0' if it is not.
-"""
-
-__question_format_multiple_choice = """
-Which sentence is more grammatical and native-like? 
-1) {sentence_1} 
-2) {sentence_2}
-"""
-
-__question_format_binary = """
-Is this sentence grammatically correct? 
-1) {sentence_1}
-"""
-
-
-def load_LLM_messages_multiple_choice(grammar_data, n_shot=5, randomize=False):
-
-    messages = [
-        {
-            "role": "system",
-            "content": __system_message_multiple_choice
-        }
-    ]
-    if randomize:
-        shuffle(grammar_data)
-    for idx in range(n_shot):
-        good_sentence = grammar_data[idx]['sentence_good']
-        bad_sentence = grammar_data[idx]['sentence_bad']
-        messages.extend([
-            {
-                "role": "user",
-                "content": __question_format_multiple_choice.format(
-                    sentence_1=good_sentence, 
-                    sentence_2=bad_sentence
-                )
-            },
-            {
-                "role": "assistant",
-                "content": '1'
-            },
-            {
-                "role": "user",
-                "content": __question_format_multiple_choice.format(
-                    sentence_1=bad_sentence, 
-                    sentence_2=good_sentence
-                )
-            },
-            {
-                "role": "assistant",
-                "content": '2'
-            }
-        ])
-    return messages
-
-
-def load_LLM_messages_binary(grammar_data, n_shot=5, randomize=False):
-    messages = [
-        {
-            "role": "system",
-            "content": __system_message_binary
-        }
-    ]
-    if randomize:
-        shuffle(grammar_data)
-    for idx in range(n_shot):
[…]

### 2025-04-22 src/util/hidden.py

@@ -1,60 +1,0 @@
-from util.data import load_grammar_data
-from util.model import clear_cuda_cache
-from tqdm import tqdm
-import numpy as np
-
-def load_hidden_states(file_path, model, tokenizer, device, batch_size=50):
-
-    grammar_data = load_grammar_data(file_path)
-    hidden_states = []
-    
-    for i in tqdm(range(0, len(grammar_data), batch_size)):
-
-        batch = grammar_data[i:i+batch_size]
-        good_sentences = [example['sentence_good'] for example in batch]
-        bad_sentences = [example['sentence_bad'] for example in batch]
-
-        good_tokens = tokenizer(good_sentences, return_tensors='pt', padding=True).to(device)
-        bad_tokens = tokenizer(bad_sentences, return_tensors='pt', padding=True).to(device)
-
-        good_tokens = {k: v.to(device) for k, v in good_tokens.items()}
-        bad_tokens = {k: v.to(device) for k, v in bad_tokens.items()}
-
-        good_outputs = model(
-            **good_tokens, 
-            output_hidden_states=True, 
-            output_attentions=True,
-            return_dict=True
-        )
-        bad_outputs = model(
-            **bad_tokens, 
-            output_hidden_states=True, 
-            output_attentions=True, 
-            return_dict=True
-        )
-
-        good_hidden_states = [v.detach().cpu().numpy() for v in good_outputs.hidden_states]
-        bad_hidden_states = [v.detach().cpu().numpy() for v in bad_outputs.hidden_states]
-
-        good_hidden_states = np.swapaxes(good_hidden_states, 0, 1)
-        bad_hidden_states = np.swapaxes(bad_hidden_states, 0, 1)
-        
-        good_attentions = [v.detach().cpu().numpy() for v in good_outputs.attentions]
-        bad_attentions = [v.detach().cpu().numpy() for v in bad_outputs.attentions]
-
-        good_attentions = np.swapaxes(good_attentions, 0, 1)
-        bad_attentions = np.swapaxes(bad_attentions, 0, 1)
-
-        for i in range(len(batch)):
-            hidden_states.append({
-                'good_sentence': good_sentences[i],
-                'bad_sentence': bad_sentences[i],
-                'good_hidden_states': good_hidden_states[i],
-                'bad_hidden_states': bad_hidden_states[i],
-                'good_attention_states': good_attentions[i],
-                'bad_attention_states': bad_attentions[i],
-            })
-
-        clear_cuda_cache()
-
-    return hidden_states

### 2025-05-07 paper/main.tex
updates

--- 2025-02-27 paper/main.tex
@@
 \usepackage{xcolor}         % colors
+\usepackage{graphicx}       % for images
+\usepackage{adjustbox}      % for figure positioning
+\usepackage{amsmath}        % for align environment
+\usepackage{mathtools}      % for \coloneqq, :=, etc.
+\usepackage{enumitem}      % customised lists (needed for [label=..., itemsep=...] syntax)
+\setlist[enumerate]{label=(\arabic*), leftmargin=*, itemsep=2pt}
+% --- extras for algorithm and example boxes ---
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
+\usepackage{mdframed}
+\usepackage{setspace}
 
-\title{Unveiling the Syntax Within: Interpreting Grammar Embeddings in Meta’s LLaMA Models}
+% lightly‑coloured framed environment for data illustrations
+\newmdenv[
+  backgroundcolor=gray!8,
+  roundcorner=4pt,
+  skipabove=\topsep,
+  skipbelow=\topsep,
+  innertopmargin=6pt,
+  innerbottommargin=6pt,
+  innerleftmargin=6pt,
+  innerrightmargin=6pt
+]{examplebox}
 
-% The \author macro works with any number of authors. There are two commands
-% used to separate the names and addresses of multiple authors: \And and \AND.
-%
-% Using \And between authors leaves it to LaTeX to determine where to break the
-% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
-% authors names on the first line, and the last on the second line, try using
-% \AND instead of \And before the third author name.
+\title{Unveiling Linguistic and Mathematical Knowledge: Interpreting Grammar and Arithmetic Embeddings in Small-Scale LLMs}
 
@@
 \begin{abstract}
-  This paper investigates the mechanisms by which large language models (LLMs) encode 
-  grammatical knowledge, focusing on Meta's LLaMA models. By leveraging embedding vectors, 
-  we classify grammatically correct sentences and analyze the activation patterns of 
-  attention heads to identify their roles in processing specific grammatical structures. 
-  Furthermore, we explore the effects of selectively removing these attention heads, shedding 
-  light on how grammar is embedded within the model's architecture. Our findings aim to enhance
-  the understanding of LLMs' linguistic capabilities and their internal organization of syntactic knowledge.
+Small transformer language models ($\leq$10 B parameters) already solve a surprising range of grammatical and numerical tasks.
+But \emph{which} internal components drive each capability—and how much circuitry is reused across domains—remains unclear.
+We study three task families—synthetic arithmetic verification, arithmetic word-problems, and grammatical acceptability—and trace responsibility down to the level of individual attention heads.
+Using a causal ablation-and-pruning procedure that extracts a \emph{Minimum-Sufficient Head Circuit} (MSHC) for each task, we show that only 10-20 heads (0.4 \% of parameters) are needed to recover 90 \% of full-model accuracy in the models we analyse (Gemma-9B, Llama-8B, Qwen-8B).
+The MSHCs for arithmetic verification and word-problems overlap by 40-60 \%, revealing a reusable numerical sub-network, whereas grammar circuits are largely disjoint.
+These findings suggest that small LLMs learn a dedicated "number sense" circuit that generalises from bare arithmetic to text-framed problems, while syntactic competence is carried by a separate set of heads.
+Our results offer new levers for parameter-efficient fine-tuning and mechanistic interpretability of compact language models.
 \end{abstract}
@@
 
-\section{Related Work}
+Large language models (LLMs) have transformed natural language processing and mathematical reasoning, demonstrating unprecedented capabilities across a spectrum of tasks—from simple question answering to complex linguistic analysis, arithmetic problem solving, and formal reasoning \citep{brown2020language, chowdhery2022palm, touvron2023llama2, jiang2023mistral}.
+These models, trained through self-supervised learning on vast corpora of text and mathematical data, have increasingly approached human-like performance in generating both grammatically well-formed text and mathematically coherent solutions, despite having no explicit grammatical rules or arithmetic algorithms programmed into their architecture.
+This emergent dual competence—arising purely from statistical patterns in training data—represents a fascinating case study in how both linguistic and mathematical knowledge can be acquired implicitly through exposure rather than explicit instruction.
 
-\begin{enumerate}
+In this work we move beyond aggregate performance metrics and zoom in on the level of individual attention heads.
+Inspired by circuit-based interpretability analyses \citep{olah2020zoom, elhage2021superposition}, we introduce the \emph{Minimum Sufficient Head Circuit} (MSHC)—the smallest set of heads whose activations suffice to solve a task within a user‑specified tolerance.
+Extracting MSHCs for arithmetic verification, mathematical word‑problems, and grammatical acceptability across three open‑weight models (Gemma‑9B, Llama‑8B, Qwen‑8B), we uncover a substantial overlap between the arithmetic and word‑problem circuits and only minimal intersection with the grammar circuit.
+These results indicate that small LLMs reuse a shared numerical sub‑circuit across superficially different tasks while maintaining a distinct pathway for syntactic reasoning.
+We quantify this overlap, analyse how it scales with model size, and discuss implications for parameter‑efficient fine‑tuning and model editing.
 
-  \item \textbf{Deciphering Stereotypes in Pre-Trained Language Models} 
-  \begin{itemize}
-    \item This paper goes into details about how stereotypes are encoded in the embeddings of pre-trained language models and how specific attention heads are responsible for encoding them.
+The field has witnessed exponential growth in model size, from early transformer models with hundreds of millions of parameters \citep{vaswani2017attention} to modern giants like GPT-4 \citep{openai2023gpt4} that likely contain trillions of parameters.
+While these massive models have captured headlines with their impressive capabilities in both language and mathematics, a parallel revolution has been unfolding in the development of smaller, more efficient models in the 1-8 billion parameter range.
+Models like Llama 2 \citep{touvron2023llama2}, Gemma 7B \citep{jiang2023mistral}, and Qwen \citep{bai2023qwen} have demonstrated remarkable performance in both linguistic and numerical tasks despite their relatively modest size, making them particularly valuable for practical applications where computational efficiency and deployment costs are significant concerns.
+
+These smaller LLMs offer a compelling balance between capability and efficiency, enabling deployment on consumer hardware, edge devices, and resource-constrained environments.
+Their reduced inference costs make them attractive for commercial applications, while their smaller memory footprint allows for fine-tuning and adaptation with more modest computational resources.
+Understanding how these models encode and represent both grammatical and numerical knowledge is therefore not merely an academic exercise but has significant practical implications for developing more capable, efficient, and cognitively robust language technologies that can handle both linguistic and mathematical reasoning.
+
+\section{Related Work and Background}
+
+\subsection{Linguistic and Mathematical Evaluation of Language Models}
+Research on evaluating neural language models has evolved from early work on LSTM-based architectures \citep{linzen2016assessing} to comprehensive assessment of transformer-based models across both linguistic and mathematical domains \citep{goldberg2019assessing, devlin2019bert, saxton2019analysing}.
+For grammatical evaluation, frameworks have progressed from simple agreement tests to comprehensive benchmarks like CoLA \citep{warstadt2019neural} and BLiMP \citep{warstadt2020blimp}.
+For mathematical reasoning, benchmarks include GSM8K \citep{cobbe2021training}, SVAMP \citep{patel2021nlpforml}, and MathQA \citep{amini2019mathqa}, which test arithmetic computation, word problem understanding, and quantitative reasoning.
+Recent studies have revealed that while performance in both domains scales with size, challenges remain with complex hierarchical structures in grammar and multi-step reasoning in mathematics \citep{thrush2022winograd, qian2022limitations, lewkowycz2022solving}.
+
+\subsection{Scale and Cognitive Competence}
+The relationship between model scale and cognitive abilities follows complex patterns beyond the general power-law scaling observed in language modeling \citep{kaplan2020scaling}.
+Research suggests that rare grammatical constructions and complex arithmetic operations require disproportionately more training data \citep{wei2021frequency, patel2021nlpforml}.
+Certain phenomena in both domains show nonlinear improvements at specific parameter thresholds \citep{zhang2023language}, with mathematical reasoning often exhibiting steeper scaling curves than linguistic tasks \citep{rae2021scaling}.
+Studies on compositional generalization indicate that scaling alone may not capture human-like cognitive productivity without architectural innovations that support both symbolic and statistical processing \citep{hu2020systematic, lake2022human}.
+
+\subsection{Probing Language Models for Cognitive Knowledge}
+Researchers have developed various probing techniques to understand how linguistic and mathematical knowledge is represented within model parameters.
+Structural probes have revealed that models implicitly encode both syntactic hierarchies and numerical relationships \citep{hewitt2019structural, wallace2019nlp}, with different types of knowledge appearing at different network depths \citep{tenney2019bert, kim2020interpreting}.
+Studies show that transformer-based models exhibit emergent capabilities resembling both discrete linguistic rules and arithmetic algorithms \citep{manning2020emergent, power2022moresymbolic}, with individual neurons specializing in specific linguistic features or numerical operations \citep{geva2021transformer, patel2022mapping}.
+Attention patterns often correspond to both syntactic dependencies and mathematical relations, with different attention heads specializing in distinct cognitive phenomena \citep{clark2019does, kim2020analyzing}.
+
+\subsection{Small LLMs and Comparative Studies}
+The proliferation of open-weight models like OPT \citep{zhang2022opt}, Llama \citep{touvron2023llama}, Gemma \citep{team2024gemma}, and Qwen \citep{bai2023qwen} has enabled more systematic analyses of how capabilities in both linguistic and mathematical domains scale and how architectural choices affect performance.
+Studies show that smaller models with innovative architectures can outperform larger ones in specific aspects of language and mathematics \citep{mckenzie2023inverse}, and data quality may be as important as scale for robust cognitive representations \citep{hoffmann2022empirical}.
+Comparative analyses across architectures remain limited but suggest significant variations in performance even at similar parameter scales, with some models showing domain-specific strengths \citep{talmor2020olmpics, zhao2023survey}.
+Recent frameworks for quantifying evaluation uncertainties \citep{roberts2023quantifying} and structured evaluation approaches \citep{xia2023structured} have revealed that architectural design choices impact both grammatical and mathematical phenomena differently, suggesting that model architectures encode cognitive knowledge in fundamentally different ways.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Methodology}
+\label{sec:methods}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Our analysis proceeds in three strands:
+
+\begin{itemize}[leftmargin=1.3em,itemsep=0.35\baselineskip]
+  \item[\textbf{(1)}]   \textbf{Low-dimensional linear separability (LS).}\;
+                       We measure how well a \emph{D-dimensional} projection
+                       of the hidden state at each layer separates correct from
+                       incorrect examples (D $\leq$ 3).
+  \item[\textbf{(2)}]   \textbf{Minimum Sufficient Head Circuit (MSHC).}\;
+                       Guided by the LS curves we isolate the smallest set of
+                       attention heads whose activations suffice for the task.
+  \item[\textbf{(3)}]   \textbf{Controlled datasets.}\;
+                       All tasks are cast as \emph{minimal pairs} so that a
+                       single factual change flips the label, letting us
+                       attribute separability (or its absence) to that fact
+                       alone.
 \end{itemize}
-  \item \textbf{}
 
+\vspace{-0.25\baselineskip}
+
+Throughout, let $\mathcal{V}$ denote the vocabulary and  
+$m:\mathcal{V}^*\!\to\!\mathbb{R}^d$ a causal transformer with $L$ layers and
+$H$ heads per layer.  For a sequence
+$x=(x_1,\dots,x_n)$ we write $\mathbf h_{x,\ell}\in\mathbb{R}^{d}$ for the activation at
+the end‑of‑sequence (EOS) token at layer $\ell$, and
+$\mathbf a_{x,\ell,h}\in\mathbb{R}^{d}$ for the contribution of
+head $h\in\{1,\dots,H\}$ in that layer.
+
+\paragraph{Activation collection.}
+For every item we construct the full prompt by concatenating a single
+one‑shot demonstration, the query sentence or equation, and the
+end‑of‑sequence marker \texttt{\textless{}/s\textgreater{}} (or its
+model‑specific analogue).  The model is executed in teacher‑forcing
+mode.  At each layer $\ell\in\{0,\dots,L\}$ we record the hidden state
+$\mathbf h_{x,\ell}$ at the position of that final EOS token—thus
+collecting a compact $L{+}1$‑vector trace that summarises the entire
+computation leading to the model’s output logit.  We omit intermediate
+attention projections and all key/value tensors to minimise I/O
+overhead; subsequent probes operate solely on this last‑token trace.
+Because all prompts fit within the context window, no padding is
+required, and teacher forcing guarantees determinism across repeated
+runs.
+
+% --- DATASETS ---
+\subsection{Task families and evaluation sets}
+\label{subsec:tasks}
+
+We consider three evaluation corpora, each built from \emph{minimal pairs}
+$(x_c,x_i)$ in which the two members differ in exactly one fact that determines
+correctness.  Every item appears in three prompt variants: the raw text, a
+two‑choice question (“A” or “B”), and a single‑candidate acceptability probe,
+each preceded by a one‑shot demonstration so that prompting remains uniform
+across models.
+
+% --- Beautified dataset examples ---
+\paragraph{Grammar \,(G).}
+The 67k sentence pairs from BLiMP \citep{warstadt2020blimp} cover twelve
+syntactic phenomena.  
+\begin{examplebox}\footnotesize
+\textbf{Good Sentence:} \textit{Who should Derek hug after shocking Richard?}\\[2pt]
+\textbf{Bad Sentence:} \textit{Who should Derek hug Richard after shocking?}
+\end{examplebox}
+
+\paragraph{Arithmetic verification \,(A).}
+We generate $10^3$ addition pairs with addends
+$n_1,n_2\in[1,10^3]$; incorrect results satisfy the deviation constraint in
+Eq.\,\eqref{eq:arith-constraint}.
+\begin{examplebox}\footnotesize
+\textbf{Good Equation:} 1338+ 88 = 1426\\[2pt]
+\textbf{Bad Equation:} 1338 + 88 = 2139
+\end{examplebox}
+
+\begin{equation}
+    \label{eq:arith-constraint}
+    .5 * (n1 \pm n2) \leq \text{noisy} (n1 \pm n2) \leq 1.5 * (n1 \pm n2)
+\end{equation}
+
+\paragraph{Word-problem arithmetic \,(W).}
+The 100 story problems are produced by numerically perturbing the ending of a
+template narrative:
+\begin{examplebox}\footnotesize
+\textbf{Good Expression:} Tim has 5 apples and eats 2, leaving him with 3 apples.\\[2pt]
+\textbf{Bad Expression:} Tim has 5 apples and eats 2, leaving him with 10 apples.
+\end{examplebox}
+where the perturbation is applied by taking a sentence generated with a template and replacing the numbers with perturbed numbers from 
+the 10000-item dataset of addition/subtraction pairs generated as in~(\ref{eq:arith-constraint}).
+% --- PCA‑SVM ---
+
+%------------------------------------------------------------------------
+\subsection{Low‑dimensional linear separability metric}
+\label{subsec:lsmetric}
+%------------------------------------------------------------------------
+
+A well‑known pathology of very high–dimensional spaces is that almost
+\emph{any} two finite clouds are linearly separable with overwhelming
+probability \citep[see][]{cover1965geometrical}.  
+Raw accuracy of a linear probe in the full residual space
+$\mathbb R^{d}$ therefore over‑states how “easy’’ a task is.  
+To obtain a more conservative—and hence more informative—measure of
+representational structure, we evaluate separability after first
+collapsing the hidden states onto just \emph{two} principal directions.
+We call the resulting statistic the
+\emph{low‑dimensional linear separability} score,
+$\operatorname{LS}_{t,\ell}$.
+
+\vspace{0.5\baselineskip}
+\noindent
+\textbf{Notation.}\;
+Fix a task $t$ and a transformer layer $\ell$.
+For every prompt $x$ we write
+$
+  \mathbf h_{x,\ell}\!\in\!\mathbb R^{d}
+$
+for the centred hidden state at the EOS position:
+$\mathbf h_{x,\ell}\;:=\;\mathbf h_{x,\ell}^{(n)}-\bar{\mathbf h}_{\ell}$,
+where
+$
+  \bar{\mathbf h}_{\ell}
+  =\tfrac1{|\mathcal D_t|}\sum_{x\in\mathcal D_t}\mathbf h_{x,\ell}^{(n)}
+$
+is the layer mean computed over the complete training split
+$\mathcal D_t$.
+
+\paragraph{Step A: variance‑maximising projection.}
+Let
+$
+  \Sigma_\ell
+  =\tfrac1{|\mathcal D_t|}
+   \sum_{x\in\mathcal D_t}\mathbf h_{x,\ell}\mathbf h_{x,\ell}^{\!\top}
+$
+be the empirical covariance matrix.
+We seek an orthonormal matrix
+$\mathbf W_\ell\!\in\!\mathbb R^{d\times 2}$ that captures the greatest
+possible variance under a rank‑2 constraint:
+\begin{equation}
+\label{eq:pca}
+\max_{\mathbf W^\top \!\mathbf W=\mathbf I_2}\;
+\operatorname{Tr}\!\bigl(\mathbf W^\top\Sigma_\ell\mathbf W\bigr).
+\end{equation}
+The optimal columns are the two leading eigenvectors of $\Sigma_\ell$.
+The corresponding 2‑D coordinates are
+$
+  \tilde{\mathbf h}_{x,\ell}
+  \;=\;
+  \mathbf W_\ell^\top\mathbf h_{x,\ell}
+  \in\mathbb R^{2}.
+$
+
+\paragraph{Step B: linear decision boundary.}
+Assign labels
+$y_i\!=\!+1$ for \emph{correct} members of a minimal pair and
+$y_i\!=\!-1$ for \emph{incorrect} ones.
+We train a soft‑margin support‑vector machine in the projected space:
+\begin{equation}
+\label{eq:svm}
+\min_{\mathbf w,b}\;
+\tfrac12\|\mathbf w\|_2^2
++\;C\sum_{i=1}^{N_t}
+\max\!\bigl(0,\,1-y_i(\mathbf w^\top\tilde{\mathbf h}_{x_i,\ell}+b)\bigr),
+\qquad
+C=10.
+\end{equation}
+Because $\tilde{\mathbf h}_{x,\ell}\!\in\!\mathbb R^2$, the resulting
+classifier depends on at most three parameters, precluding the kind of
+pathological over‑fitting that haunts full‑dimensional probes.
+
+\paragraph{Definition of the metric.}
+Let
+$
+  \hat y_i
+  =\operatorname{sign}\!\bigl(\mathbf w^\top\tilde{\mathbf h}_{x_i,\ell}+b\bigr)
+$
+denote the SVM’s prediction on a held‑out validation example.
+We define
+\begin{equation}
+\label{eq:ls-score}
+\operatorname{LS}_{t,\ell}
+=\frac{1}{N_t^{\text{val}}}
+ \sum_{i=1}^{N_t^{\text{val}}}\mathbf 1[\hat y_i=y_i],
+\qquad
+\operatorname{LS}_{t,\ell}\in[0,1].
+\end{equation}
+
+\paragraph{Interpretation.}
+Random guessing gives $\operatorname{LS}_{t,\ell}=0.5$.
+A score close to 1 implies that the first two principal axes already
+support a linear decision boundary, i.e.\ the class‑conditional embeddings
+differ in a \emph{low‑codimension} direction.
+Conversely, scores near 0.5 indicate either geometric entanglement or
+dispersion of the signal across many dimensions.
+
+\vspace{0.3\baselineskip}
+In what follows we plot $\ell\mapsto\operatorname{LS}_{t,\ell}$ for each
+task. Peaks in these curves spotlight layers where the model’s hidden
+states make the good/bad distinction in the simplest possible way: by
+shifting along just two orthogonal directions.
+
+% ---------- NEW MSHC SUBSECTION ----------
+\subsection{The Minimum Sufficient Head Circuit}
+\label{subsec:mshc}
+
+A transformer’s computation is frequently dominated by a surprisingly
+small subset of its attention heads.  We formalise this with the
+\emph{Minimum Sufficient Head Circuit} (MSHC).  Fix an accuracy tolerance
+$\epsilon\in(0,0.5)$.  Let
+$\mathcal H=\{(\ell,h):1\le\ell\le L,\;1\le h\le H\}$ be the set of all
+heads.  A subset $\mathcal C\subseteq\mathcal H$ is an
+\textbf{MSHC}${}_{\!\epsilon}$ if, \emph{with high probability (WHP)}, enabling any single head from $\mathcal C$ is already sufficient to lift accuracy above chance:
+\[
+  (\forall h\in\mathcal C)\quad
+  \Pr_{x\sim\mathcal D}\!\bigl[\mathrm{Acc}\bigl(m;\{h\}\bigr) > 0.5+\epsilon\bigr] \;\ge\; 1-\delta,
+\]
+where $\delta$ is a user‑specified failure probability (we use $\delta=0.05$ in all experiments), and $\mathcal C$ is inclusion‑minimal with this property.
+
+%
+\noindent\textbf{Hunting the circuit}
+Our search still employs a sliding window but now ranks layers by the \emph{size} of the accuracy loss they incur.  
+We slide a window of width $w=\lfloor xL\rfloor$ layers from the bottom to the top of the network, disable \emph{all} heads in that window, and measure the resulting accuracy $\mathrm{Acc}_{\mathrm{off}}$.
+For every layer $\ell$ covered by the current window we store  
+$\textsc{AccLayer}[\ell]=\max\!\bigl(\textsc{AccLayer}[\ell],\mathrm{Acc}_{\mathrm{off}}\bigr)$—the \emph{best} accuracy ever observed when \emph{any} window that contains $\ell$ is switched off.  
+After scanning all windows we compute a per‑layer drop score  
+$\Delta_\ell=\mathrm{Acc}_{\mathrm{full}}-\textsc{AccLayer}[\ell]$  
+and keep those layers whose $\Delta_\ell$ lies in the top 25th percentile.  
+These high‑impact layers seed the head‑level search:
+
+% --- begin patch: enumerate steps for MSHC ---
+\begin{enumerate}[label=\textbf{Step \arabic*.}, leftmargin=4em, itemsep=0pt]
+\item Disable every head \emph{in the selected top‑quartile layers}, establishing a ``dark‑start'' baseline.
+\item Run a stochastic pruning loop (Alg.\,\ref{alg:mshc}) that begins with bundle size $k_0=\lceil 2\sqrt{|\mathcal C|}\rceil$. For each bundle size $k$, we draw $N$ random bundles of $k$ heads, measure their accuracies, and remove the bundle that attains the \emph{lowest} accuracy whenever that value is $\le 0.5+\epsilon$. When the worst of the $N$ bundles exceeds the threshold, we tighten the constraint by updating $k\leftarrow\max(1,\lfloor k/2\rfloor)$ and continue.
+\item Stop when $k=1$ and every random bundle succeeds—by definition, the remaining heads form an MSHC.
 \end{enumerate}
+% --- end patch: enumerate steps for MSHC ---
+The procedure terminates because the candidate set shrinks
+monotonically and can be pruned at most $|\mathcal H|$ times.
 
-\section{Methodology}
+The narrative interpretation is simple: we first find \emph{where} knowledge
+lives, then keep only those filaments that reliably relight the lamp.
+\begin{algorithm}[t]
+\setstretch{1.15}
+\caption{Sliding‑window percentile localisation followed by stochastic discovery of an MSHC${}_{\epsilon}$}
+\label{alg:mshc}
+\begin{algorithmic}[1]
+\Require Window fraction $x$, tolerance $\epsilon$, sample count $N$
+\State $\mathrm{Acc}_{\mathrm{full}}\gets\mathrm{Acc}(m)$
+\State $w\gets\lfloor xL\rfloor$
+\State initialise array $\textsc{AccLayer}[1\!:\!L]\gets 0$ \Comment{best accuracy seen with each layer disabled}
+\For{$s = 1$ \textbf{to} $L-w+1$}
+    \State disable all heads in layers $s$ to $s+w-1$
+    \State $\mathrm{Acc}_{\mathrm{off}}\gets\mathrm{Acc}(m)$
+    \For{$\ell = s$ \textbf{to} $s+w-1$}
+        \State $\textsc{AccLayer}[\ell]\gets\max\!\bigl(\textsc{AccLayer}[\ell],\,\mathrm{Acc}_{\mathrm{off}}\bigr)$
+    \EndFor
+    \State re‑enable the disabled layers
+\EndFor
+\State \Comment{compute per‑layer drop scores}
+\For{$\ell = 1$ \textbf{to} $L$}
+    \State $\textsc{Drop}[\ell]\gets\mathrm{Acc}_{\mathrm{full}} - \textsc{AccLayer}[\ell]$
+\EndFor
+\State $\tau\gets\text{75th percentile of }\textsc{Drop}$
+\State $\mathcal{L}\gets\{\ell\;|\;\textsc{Drop}[\ell]\ge\tau\}$ \Comment{top‑quartile layers}
+\State $\mathcal{C}\gets$ all heads in $\mathcal{L}$ \Comment{initial candidate circuit}
+\State $k\gets\left\lceil 2\sqrt{|\mathcal C|}\right\rceil$
+\While{$k \ge 1$}
+    \Repeat
+        \State $\text{minAcc}\gets 1$; $\mathcal K_{\mathrm{min}}\gets\emptyset$
+        \For{$i = 1$ \textbf{to} $N$}
+            \State draw $\mathcal K\!\sim\!\mathrm{Unif}\{\mathcal S\subseteq\mathcal C:|\mathcal S|=k\}$
+            \State $\text{acc}\gets\mathrm{Acc}(m;\mathcal K)$
+            \If{$\text{acc} < \text{minAcc}$}
+                \State $\text{minAcc}\gets\text{acc}$; $\mathcal K_{\mathrm{min}}\gets\mathcal K$
+            \EndIf
+        \EndFor
+        \If{$\text{minAcc}\le 0.5+\epsilon$}
+            \State $\mathcal C\gets\mathcal C\setminus\mathcal K_{\mathrm{min}}$ \Comment{prune the worst bundle}
+        \EndIf
+    \Until{$\text{minAcc}>0.5+\epsilon$}
+    \State $k\gets\max(1,\lfloor k/2\rfloor)$
+\EndWhile
+\State\Return $\mathcal C$
+\end{algorithmic}
+\end{algorithm}
 
+Empirically, the circuit coalesces in fewer than 300 iterations on all three
+models studied—Gemma-9B, Llama-8B, and Qwen-8B.
+
+% ---------- NEW PROTOCOL PARAGRAPH ----------
+\subsection{Experimental protocol}
+\label{subsec:protocol}
+
+Our analysis focuses on three open-weight checkpoints that occupy a comparable
+size class yet differ architecturally: \emph{Gemma-9B}, \emph{Llama-8B} and
+\emph{Qwen-8B}.  Each model is evaluated on exactly the same train/validation/
+test splits (60 : 20 : 20) of the G, A, and W corpora described above.  For a
+given model we first trace the accuracy‑vs‑layer curves
+(§\ref{subsec:pca_svm}), then feed the validation set to the sliding-window
+search with $(\epsilon,k)=(0.1,3)$ and a window fraction $x\!=\!0.2$.  The
+resulting MSHC is frozen before we inspect test performance, and uncertainty
+bands are computed via 1 000-sample bootstrap over minimal pairs.  All
+experiments run on a single A100 80 GB GPU; wall-clock times are listed in
+Appendix C.
 \section{Experiments}

### 2025-05-07 src/grammar.py

@@ -1,0 +1,12 @@
+from util.data import load_combined_grammar_data
+
+grammar_data = load_combined_grammar_data(n_sample=1000)
+
+fields = set()
+linguistics_terms = set()
+for example in grammar_data:
+    fields.add(example["field"])
+    linguistics_terms.add(example["linguistics_term"])
+
+print(fields)
+print(linguistics_terms)

### 2025-05-07 src/module/algo.py

@@ -1,0 +1,209 @@
+import torch
+import numpy as np
+from util.model import clear_cuda_cache, get_device_memory_report, load_model
+from module.blocked_attention import (
+    load_altered_attention_model,
+    disable_head,
+    enable_head,
+    get_model_dimensions,
+    get_disabled_heads
+)
+from util.chat import load_embeddings_dataset, load_embeddings_dataset_batch, load_attentions_dataset
+from transformers import AutoModelForCausalLM, AutoTokenizer
+from sklearn.svm import SVC
+from sklearn.metrics import accuracy_score
+from sklearn.decomposition import PCA
+from tqdm import tqdm
+from random import sample
+import json
+from util.data import load_numerical_data
+from random import shuffle
+from pprint import pprint
+
+
+def metric(
+    data,
+    model,
+    tokenizer,
+    device,
+    metric_type: str,
+    problem_type: str,
+    use_system_message: bool = False,
+    D: int = 3,
+):
+    embeddings = load_embeddings_dataset(
+        data,
+        model,
+        tokenizer,
+        device,
+        types_=[metric_type],
+        use_tqdm=True,
+        problem_type=problem_type,
+        use_system_message=use_system_message,
+    )[metric_type]
+    X = np.array([embedding["states"][-1] for embedding in embeddings])
+    y = np.array([embedding["answer"] for embedding in embeddings])
+
+    pca = PCA(n_components=D)
+    pca.fit(X)
+    X_pca = pca.transform(X)
+
+    svm = SVC(kernel="linear")
+    svm.fit(X_pca, y)
+    acc = accuracy_score(y, svm.predict(X_pca))
+    return acc
+
+
+def disable_layers(model, layer_subset, n_heads):
+    for layer in layer_subset:
+        for head in range(n_heads):
+            disable_head(model, layer, head)
+
+
+def enable_layers(model, layer_subset, n_heads):
+    for layer in layer_subset:
+        for head in range(n_heads):
+            enable_head(model, layer, head)
+
+
+def layer_select(
+    data,
+    model,
+    tokenizer,
+    device,
+    metric_type="raw",
+    problem_type="numerical",
+    use_system_message=False,
+    percentages=[1/2, 1/3, 1/4, 1/5, 1/8, 1/10, 1/20],
+    outfile="outputs/removed_layers_numerical.json",
+):
+    n_layers, n_heads = get_model_dimensions(model, tokenizer, device)
+    layer_pool = [l for l in range(n_layers)]
+    results = {}
[…]

### 2025-05-07 src/module/blocked_attention.py

@@ -11,7 +11,7 @@ from transformers.models.gemma2.modeling_gemma2 import (
     Gemma2Attention,
 )
 import numpy as np
-from typing import Tuple, Optional, Unpack, Callable, Dict, List
+from typing import Tuple, Optional, Unpack, Callable, Dict, List, Set
 import torch
 import torch.nn as nn
 from util.model import clear_cuda_cache
@@ -98,17 +98,17 @@ class GemmaBlockedAttention(Gemma2Attention):
         self,
         config,
         layer_idx,
-        disable_idx: Optional[List[int]] = [],
+        disable_idx: Optional[Set[int]] = {},
         disable_positional_encoding: Optional[bool] = False,
     ):
         super().__init__(config, layer_idx)
-        self.disable_idx = disable_idx
+        self.disable_idx = set(disable_idx)
         self.disable_positional_encoding = disable_positional_encoding
         self.save_model_dimensions = True if layer_idx == 0 else False
         self.model_dimensions = None
 
     def disable_head(self, head_idx: int):
-        self.disable_idx.append(head_idx)
+        self.disable_idx.add(head_idx)
 
     def enable_head(self, head_idx: int):
         if head_idx in self.disable_idx:
@@ -189,8 +189,8 @@ class GemmaBlockedAttention(Gemma2Attention):
         )
 
         if self.disable_idx:
-            # print(self.disable_idx)
-            attn_output[:, :, np.array(self.disable_idx), :] = 0
+            disable_idx = np.array(list(self.disable_idx))
+            attn_output[:, :, disable_idx, :] = 0
 
         if self.save_model_dimensions:
             self.model_dimensions = attn_output.shape[-2]
@@ -205,9 +205,9 @@ class LlamaBlockedAttention(LlamaAttention):
         self,
         config,
         layer_idx,
-        disable_idx: Optional[List[int]] = [],
-        raise_entropy_idx: Optional[List[int]] = [],
-        lower_entropy_idx: Optional[List[int]] = [],
+        disable_idx: Optional[Set[int]] = {},
+        raise_entropy_idx: Optional[Set[int]] = {},
+        lower_entropy_idx: Optional[Set[int]] = {},
         disable_positional_encoding: Optional[bool] = False,
     ):
         super().__init__(config, layer_idx)
@@ -219,7 +219,7 @@ class LlamaBlockedAttention(LlamaAttention):
         self.model_dimensions = None
 
[…]

### 2025-05-07 src/module/heads.py

@@ -1,0 +1,181 @@
+import torch
+import numpy as np
+from util.model import clear_cuda_cache, get_device_memory_report, load_model
+from module.blocked_attention import (
+    load_altered_attention_model,
+    disable_head,
+    enable_head,
+    get_model_dimensions,
+    get_disabled_heads
+)
+from util.chat import load_embeddings_dataset, load_embeddings_dataset_batch, load_attentions_dataset
+from transformers import AutoModelForCausalLM, AutoTokenizer
+from sklearn.svm import SVC
+from sklearn.metrics import accuracy_score
+from sklearn.decomposition import PCA
+from tqdm import tqdm
+from random import sample
+import json
+from util.data import load_numerical_data
+from random import shuffle
+from pprint import pprint
+
+
+def get_splits(layer_pool, percentage):
+    n_per_split = int(len(layer_pool) * percentage)
+    splits = []
+    for i in range(0, len(layer_pool), n_per_split):
+        if i + n_per_split > len(layer_pool):
+            splits[-1].extend(layer_pool[i:])
+        else:
+            splits.append(layer_pool[i:i+n_per_split])
+    return splits
+
+
+def metric(
+    embeddings,
+    metric_type: str,
+):
+    embeddings = embeddings[metric_type]
+    X = np.array([embedding["states"][-1] for embedding in embeddings])
+    y = np.array([embedding["answer"] for embedding in embeddings])
+
+    pca = PCA(n_components=2)
+    pca.fit(X)
+    X_pca = pca.transform(X)
+
+    svm = SVC(kernel="linear")
+    svm.fit(X_pca, y)
+    acc = accuracy_score(y, svm.predict(X_pca))
+    return acc
+
+
+def head_select(
+    data,
+    model,
+    tokenizer,
+    device,
+    metric_type="raw",
+    problem_type="numerical",
+    use_system_message=False,
+    percentages=[1/2, 1/3, 1/4, 1/5, 1/8, 1/10, 1/20],
+    outfile="removed_layers_numerical.json",
+):
+    n_layers, n_heads = get_model_dimensions(model, tokenizer, device)
+    layer_pool = [l for l in range(n_layers)]
+    results = []
+    for percentage in percentages:
+        splits = get_splits(layer_pool, percentage)
+        split_results = []
+        print('Disabling layers: ', splits)
+        for split in splits:
+            for layer in split:
+                for head in range(n_heads):
+                    disable_head(model, layer, head)
+            print('Disabling layers: ', split)
[…]

### 2025-05-07 src/t.py

@@ -1,0 +1,79 @@
+from module.blocked_attention import load_altered_attention_model
+from util.data import load_numerical_data, load_grammar_data_by_category, load_word_problems_data
+import torch
+from random import shuffle
+from module.algo import layer_select, head_select
+import json
+from pprint import pprint
+
+if __name__ == "__main__":
+    numerical_data = load_numerical_data(margin=50)
+    shuffle(numerical_data)
+    numerical_data = numerical_data[:50]
+
+    noun_agreement_data = load_grammar_data_by_category(
+        category="linguistics_term", value="determiner_noun_agreement", sample_rate=1
+    )
+    shuffle(noun_agreement_data)
+    noun_agreement_data = noun_agreement_data[:50]
+
+    word_problems_data = load_word_problems_data()
+
+    device = torch.device("cuda:6" if torch.cuda.is_available() else "cpu")
+    model, tokenizer = load_altered_attention_model(
+        "google/gemma-2-9b-it", device, model_type="gemma"
+    )
+    with open("outputs/gemma_removed_layers_numerical.json", "r") as f:
+        numerical_layer_data = json.load(f)
+
+    with open("outputs/gemma_removed_layers_word_problems.json", "r") as f:
+        word_problems_layer_data = json.load(f)
+
+    with open("outputs/gemma_removed_layers_noun_agreement.json", "r") as f:
+        noun_agreement_layer_data = json.load(f)
+
+    head_select(
+        word_problems_data,
+        model,
+        tokenizer,
+        device,
+        layer_data=word_problems_layer_data,
+        metric_type="raw",
+        problem_type="word_problems",
+        use_system_message=False,
+        outfile="outputs/gemma_removed_head_pool_word_problems.json",
+    )
+
+    # layer_select(
+    #     numerical_data,
+    #     model,
+    #     tokenizer,
+    #     device,
+    #     metric_type="raw",
+    #     problem_type="numerical",
+    #     use_system_message=False,
+    #     percentages=[1/20],
+    #     outfile="outputs/gemma_removed_layers_numerical.json",
+    # )
+    # layer_select(
+    #     word_problems_data,
+    #     model,
+    #     tokenizer,
+    #     device,
+    #     metric_type="raw",
+    #     problem_type="word_problems",
+    #     use_system_message=False,
+    #     percentages=[1/3, 1/4, 1/5, 1/8, 1/10, 1/20],
[…]

### 2025-05-07 src/test.py

@@ -1,0 +1,62 @@
+from module.heads import head_select, remove_high_entopy_heads
+from module.blocked_attention import load_altered_attention_model
+from util.data import load_numerical_data, load_grammar_data_by_category, load_word_problems_data
+import torch
+from random import shuffle
+
+
+if __name__ == "__main__":
+    numerical_data = load_numerical_data(margin=50)
+    shuffle(numerical_data)
+    numerical_data = numerical_data[:100]
+    grammar_data = load_grammar_data_by_category(
+        category="linguistics_term", value="argument_structure", sample_rate=1
+    )
+    shuffle(grammar_data)
+    grammar_data = grammar_data[:100]
+    device = torch.device("cuda:6" if torch.cuda.is_available() else "cpu")
+    model, tokenizer = load_altered_attention_model(
+        "google/gemma-2-9b-it", device, model_type="gemma"
+    )
+    word_problems_data = load_word_problems_data()
+
+    target_layers = [0, 1, 2, 3, ]
+    target_heads = {
+        0: [2, 4, 5, 6, 9, 11, 13, 14],
+        1: [1, 3, 4, 5, 8, 9, 10, 11, 12, 15],
+        2: [0, 1, 3, 4, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15],
+        3: [0, 1, 2, 3, 4, 5, 9, 11, 12],
+        27: [2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 13],
+        28: [1, 2, 4, 5, 6, 7, 8, 9, 10, 11, 14, 15],
+        29: [0, 1, 4, 5, 6, 10, 15],
+        38: [0, 2, 3, 4, 5, 6, 7, 8, 10, 11, 13, 14, 15],
+        39: [1, 2, 4, 5, 6, 10, 11, 12, 13, 14]
+    }
+    target_heads = {
+        l: [i for i in range(16)] for l in target_layers
+    }
+    
+    remove_high_entopy_heads(
+        word_problems_data,
+        model,
+        tokenizer,
+        device,
+        target_heads=target_heads,
+        metric_type="raw",
+        problem_type="word_problems",
+        percentage_to_remove=1,
+        use_system_message=False,
+        head_batch_size=10,
+        # outfile="removed_layers_numerical.json",
+    )
+    # attentions = get_head_attentions_dataset(
+    #     numerical_data,
+    #     model,
+    #     tokenizer,
+    #     device,
[…]

### 2025-05-07 src/tests.py

@@ -16,7 +16,7 @@ if __name__ == "__main__" or True:
     with open("activated.json", "w") as f:
         json.dump([(0, 0)], f)
 
-    device = torch.device("cuda:7" if torch.cuda.is_available() else "cpu")
+    device = torch.device("cuda:5" if torch.cuda.is_available() else "cpu")
     get_device_memory_report(device)
     model_name = "google/gemma-2-9b-it"
     model, tokenizer = load_altered_attention_model(model_name, device, model_type="gemma")

### 2025-05-07 src/util/chat.py

@@ -80,6 +80,46 @@ __question_format_binary_arithmetic = """Is this arithmetic expression correct?
 
 ONLY OUTPUT '1' OR '0', NOTHING ELSE."""
 
+__system_message_multiple_choice_word_problems = """You are a helpful assistant that will help me figure out which statement is correct.
+You should only output the number of the correct statement, '1' if the first statement is correct, '2' if the second statement is correct.
+
+ONLY OUTPUT THE NUMBER OF THE CORRECT SENTENCE, NOTHING ELSE.
+
+For example:
+Which statement is correct?
+1) I ate 2 apples and 2 oranges, the total number of fruits I ate is 4.
+2) I ate 2 apples and 2 oranges, the total number of fruits I ate is 5.
+
+1"""
+
+__question_format_multiple_choice_word_problems = """Which statement is correct?
+1) {sentence_1}
+2) {sentence_2}
+
+ONLY OUTPUT THE NUMBER OF THE CORRECT SENTENCE (1 OR 2), NOTHING ELSE."""
+
+__system_message_binary_word_problems = """You are a helpful assistant that will help me figure out if a statement is correct.
+You should only output '1' if the statement is correct, and '0' if it is not.
+
+ONLY OUTPUT '1' OR '0', NOTHING ELSE.
+
+For example:
+Is this statement correct?
+1) I ate 2 apples and 2 oranges, the total number of fruits I ate is 4.
+
+1"""
+
+__question_format_binary_word_problems = """Is this statement correct?
+1) {sentence_1}
+
+ONLY OUTPUT '1' OR '0', NOTHING ELSE."""
+
+__question_format_multiple_choice_word_problems = """Which statement is correct?
+1) {sentence_1}
+2) {sentence_2}
+
+ONLY OUTPUT THE NUMBER OF THE CORRECT SENTENCE (1 OR 2), NOTHING ELSE."""   
+
 
 def parse_output(output):
     if output not in ["0", "1", "2"]:
@@ -89,14 +129,17 @@ def parse_output(output):
 
 
 def load_LLM_messages_multiple_choice(
-    shots, grammar=True, randomize=False, use_system_message=True
+    shots, type_=True, randomize=False, use_system_message=True
 ):
-    if grammar:
+    if type_ == "grammar":
         system_message = __system_message_multiple_choice_grammar
         question_format = __question_format_multiple_choice_grammar
-    else:
+    elif type_ == "arithmetic":
         system_message = __system_message_multiple_choice_arithmetic
[…]

### 2025-05-07 src/util/data.py

@@ -3,7 +3,7 @@ import pickle
 import os
 import numpy as np
 from tqdm import tqdm
-
+from random import uniform
 
 def load_grammar_data(file_path):
     with open(file_path, "r") as f:
@@ -25,12 +25,7 @@ def load_combined_grammar_data(data_path: str = "../data/blimp", n_sample: int =
 
     data = []
     for example in combined_grammar_data:
-        data.append(
-            {
-                "good": example["sentence_good"],
-                "bad": example["sentence_bad"],
-            }
-        )
+        data.append(example)
     return data
 
 
@@ -61,7 +56,7 @@ def load_grammar_data_by_category(
 
 
 def load_numerical_data(
-    data_path: str = "../data/arithmetic/addition.jsonl",
+    data_path: str = "../data/arithmetic/addition_subtraction.jsonl",
     margin: int = 25,
     use_tqdm: bool = False,
 ):
@@ -75,12 +70,26 @@ def load_numerical_data(
                 "good": example["good"],
                 "bad": (
                     example[f"bad_p_{margin}"]
-                    if np.random.rand() < 0.5
+                    if uniform(0, 1) < 0.5
                     else example[f"bad_m_{margin}"]
                 ),
             }
         )
     return ret_data
+
+
+def load_word_problems_data(
+    data_path: str = "../data/word_problems/fill_in_addition_subtraction_simple.jsonl",
+    use_tqdm: bool = False,
+):
+    data = []
+    with open(data_path, "r") as f:
+        for line in f:
+            try:
+                data.append(json.loads(line))
+            except:
+                print(line)
+    return data
 
 
 def load_hidden_data(model_name, grammar_name, data_path="../data/blimp-hidden"):

### 2025-05-07 src/util/numerical.py

@@ -1,0 +1,107 @@
+from random import randint, uniform, shuffle
+import json
+
+
+def create_dataset(n_samples, digit_range=(0, 2)):
+    data = []
+    for _ in range(n_samples):
+        d1, d2 = randint(digit_range[0], digit_range[1]), randint(digit_range[0], digit_range[1])
+        a, b = randint(10 ** d1, 10 ** (d1 + 1)), randint(10 ** d2, 10 ** (d2 + 1))
+        
+        if uniform(0, 1) < 0.5:
+            ans = a + b
+            data.append({
+                'good': f'{a} + {b} = {ans}',
+                'bad_p_1': f'{a} + {b} = {ans + 1}',
+                'bad_m_1': f'{a} + {b} = {ans - 1}',
+                'bag_p_10': f'{a} + {b} = {int(ans * 1.1)}',
+                'bag_m_10': f'{a} + {b} = {int(ans * 0.9)}',
+                'bad_p_25': f'{a} + {b} = {int(ans * 1.25)}',
+                'bad_m_25': f'{a} + {b} = {int(ans * 0.75)}',
+                'bad_p_50': f'{a} + {b} = {int(ans * 1.5)}',
+                'bad_m_50': f'{a} + {b} = {int(ans * 0.5)}',
+                'bad_p_95': f'{a} + {b} = {int(ans * 1.95)}',
+                'bad_m_95': f'{a} + {b} = {int(ans * 0.05)}',
+            })
+        else:
+            a, b = max(a, b), min(a, b)
+            ans = a - b
+            data.append({
+                'good': f'{a} - {b} = {ans}',
+                'bad_p_1': f'{a} - {b} = {ans + 1}',
+                'bad_m_1': f'{a} - {b} = {ans - 1}',
+                'bag_p_10': f'{a} - {b} = {int(ans * 1.1)}',
+                'bag_m_10': f'{a} - {b} = {int(ans * 0.9)}',
+                'bad_p_25': f'{a} - {b} = {int(ans * 1.25)}',
+                'bad_m_25': f'{a} - {b} = {int(ans * 0.75)}',
+                'bad_p_50': f'{a} - {b} = {int(ans * 1.5)}',
+                'bad_m_50': f'{a} - {b} = {int(ans * 0.5)}',
+                'bad_p_95': f'{a} - {b} = {int(ans * 1.95)}',
+                'bad_m_95': f'{a} - {b} = {int(ans * 0.05)}',
+            })
+
[…]

### 2025-05-10 old2/chat.py
backup

@@ -39,12 +39,16 @@ Is this sentence grammatically correct?
 
 1"""
 
-__question_format_multiple_choice_grammar = """Which sentence is more grammatically correct? 
+__question_format_multiple_choice_grammar = """Which sentence is more grammatical and native-like? 
 1) {sentence_1} 
-2) {sentence_2}"""
+2) {sentence_2}
+
+ONLY OUTPUT THE NUMBER OF THE CORRECT SENTENCE (1 OR 2), NOTHING ELSE."""
 
 __question_format_binary_grammar = """Is this sentence grammatically correct? 
-1) {sentence_1}"""
+1) {sentence_1}
+
+ONLY OUTPUT '1' OR '0', NOTHING ELSE."""
 
 __system_message_multiple_choice_arithmetic = """You are a helpful assistant that will help me determine which of two arithmetic expressions is correct.
 
@@ -58,7 +62,9 @@ Which of the following arithmetic expressions is correct?
 
 __question_format_multiple_choice_arithmetic = """Which of the following arithmetic expressions is correct?
 1) {sentence_1}
-2) {sentence_2}"""
+2) {sentence_2}
+
+ONLY OUTPUT THE NUMBER OF THE CORRECT EXPRESSION (1 OR 2), NOTHING ELSE."""
 
 __system_message_binary_arithmetic = """You are a helpful assistant that will help me determine if an arithmetic expression is correct.
 
@@ -70,7 +76,9 @@ Is this arithmetic expression correct?
 1"""
 
 __question_format_binary_arithmetic = """Is this arithmetic expression correct?
-1) {sentence_1}"""
+1) {sentence_1}
+
+ONLY OUTPUT '1' OR '0', NOTHING ELSE."""
 
 __system_message_multiple_choice_word_problems = """You are a helpful assistant that will help me figure out which statement is correct.
 You should only output the number of the correct statement, '1' if the first statement is correct, '2' if the second statement is correct.
@@ -84,6 +92,12 @@ Which statement is correct?
 
 1"""
 
+__question_format_multiple_choice_word_problems = """Which statement is correct?
+1) {sentence_1}
+2) {sentence_2}
+
+ONLY OUTPUT THE NUMBER OF THE CORRECT SENTENCE (1 OR 2), NOTHING ELSE."""
+
 __system_message_binary_word_problems = """You are a helpful assistant that will help me figure out if a statement is correct.
 You should only output '1' if the statement is correct, and '0' if it is not.
 
[…]

### 2025-05-10 src/module/algo.py

@@ -28,8 +28,7 @@ def metric(
     device,
     metric_type: str,
     problem_type: str,
-    use_system_message: bool = False,
-    D: int = 3,
+    D: int = 10,
 ):
     embeddings = load_embeddings_dataset(
         data,
@@ -39,7 +38,6 @@ def metric(
         types_=[metric_type],
         use_tqdm=True,
         problem_type=problem_type,
-        use_system_message=use_system_message,
     )[metric_type]
     X = np.array([embedding["states"][-1] for embedding in embeddings])
     y = np.array([embedding["answer"] for embedding in embeddings])
@@ -73,29 +71,27 @@ def layer_select(
     device,
     metric_type="raw",
     problem_type="numerical",
-    use_system_message=False,
-    percentages=[1/2, 1/3, 1/4, 1/5, 1/8, 1/10, 1/20],
+    k: int = 5,
+    D: int = 10,
     outfile="outputs/removed_layers_numerical.json",
 ):
     n_layers, n_heads = get_model_dimensions(model, tokenizer, device)
     layer_pool = [l for l in range(n_layers)]
+    
     results = {}
-    for percentage in percentages:
-        results[percentage] = {}
-        k = round(len(layer_pool) * percentage)
-        print('Percentage: ', percentage, 'k: ', k)
-        for i in range(len(layer_pool) - k + 1):
-            layer_subset = layer_pool[i:i+k]
-            disable_layers(model, layer_subset, n_heads)
-            acc = metric(data, model, tokenizer, device, metric_type, problem_type, use_system_message)
-            for l in layer_subset:
-                if l not in results[percentage]:
-                    results[percentage][l] = []
-                results[percentage][l].append(acc)
-            enable_layers(model, layer_subset, n_heads)
+    for i in range(len(layer_pool) - k + 1):
+        layer_subset = layer_pool[i:i+k]
+        disable_layers(model, layer_subset, n_heads)
+        acc = metric(data, model, tokenizer, device, metric_type, problem_type, D=D)
+        for l in layer_subset:
+            if l not in results:
+                results[l] = []
+            results[l].append(acc)
+        enable_layers(model, layer_subset, n_heads)
         
         with open(outfile, "w") as f:
             json.dump(results, f, indent=4)
+    return results
 
 
 def save_head_pool(head_pool, outfile):
@@ -122,37 +118,33 @@ def head_select(
     device, 
     epsilon=.1,
     layer_data=None,
[…]

### 2025-05-10 src/module/blocked_attention.py

@@ -1,309 +1,11 @@
-from transformers import LlamaForCausalLM, AutoTokenizer, AutoModelForCausalLM
-from transformers.models.llama.modeling_llama import (
-    LlamaAttention,
-    apply_rotary_pos_emb,
-    eager_attention_forward,
-    logger,
-    ALL_ATTENTION_FUNCTIONS,
-    repeat_kv,
-)
-from transformers.models.gemma2.modeling_gemma2 import (
-    Gemma2Attention,
-)
-import numpy as np
-from typing import Tuple, Optional, Unpack, Callable, Dict, List, Set
+from module.llama import LlamaBlockedAttention
+from module.gemma import GemmaBlockedAttention
+from module.qwen import QwenBlockedAttention
+
+from transformers import AutoModelForCausalLM, AutoTokenizer
 import torch
-import torch.nn as nn
+from typing import Tuple, Dict, List
 from util.model import clear_cuda_cache
-
-from transformers.cache_utils import Cache
-from transformers.modeling_flash_attention_utils import FlashAttentionKwargs
-
-
-def print_matrix(matrix: np.ndarray):
-    for row in matrix:
-        for col in row:
-            print(f"{col:.2f}", end=" ")
-        print()
-
-
-def increase_entropy(distribution, power=2, percentile=0.2):
-    mean = torch.mean(distribution)
-    deviation = distribution - mean
-
-    over_mean = torch.clamp(torch.sign(deviation) * (deviation) ** power, 0, 1)
-    under_mean = torch.clamp(torch.sign(deviation) * (deviation) ** power, -1, 0)
-
-    over_sum = torch.sum(over_mean)
-    under_sum = torch.sum(under_mean)
-
-    sub = torch.zeros_like(distribution)
-    add = torch.zeros_like(distribution)
-
-    if over_sum > 0:
-        sub = (over_mean / over_sum) * percentile
-
-    if under_sum < 0:
-        add = (under_mean / under_sum) * percentile
-
-    new_dist = distribution - sub + add
-    return new_dist
-
-
-def increase_entropy_batch(attn_weights, power=2, percentile=0.1):
-    # Extract the last row for all samples and heads
-    last_row = attn_weights[:, :, -1, :]
-    # Calculate deviation from mean (along the token dimension)
-    mean = torch.mean(last_row, dim=-1, keepdim=True)
-    deviation = last_row - mean
-    # Calculate components above and below mean
[…]

### 2025-05-10 src/module/gemma.py

@@ -1,0 +1,121 @@
+from transformers.models.gemma2.modeling_gemma2 import (
+    Gemma2Attention,
+    eager_attention_forward,
+    apply_rotary_pos_emb,
+    ALL_ATTENTION_FUNCTIONS,
+)
+import torch
+from typing import Tuple, Optional, Unpack, Callable, Dict, List, Set
+import numpy as np
+from transformers.cache_utils import Cache
+from transformers.modeling_flash_attention_utils import FlashAttentionKwargs
+from transformers.utils import logging
+
+logger = logging.get_logger(__name__)
+
+
+class GemmaBlockedAttention(Gemma2Attention):
+    def __init__(
+        self,
+        config,
+        layer_idx,
+        disable_idx: Optional[Set[int]] = {},
+        disable_positional_encoding: Optional[bool] = False,
+    ):
+        super().__init__(config, layer_idx)
+        self.disable_idx = set(disable_idx)
+        self.disable_positional_encoding = disable_positional_encoding
+        self.save_model_dimensions = True if layer_idx == 0 else False
+        self.model_dimensions = None
+
+    def disable_head(self, head_idx: int):
+        self.disable_idx.add(head_idx)
+
+    def enable_head(self, head_idx: int):
+        if head_idx in self.disable_idx:
+            self.disable_idx.remove(head_idx)
+
+    def get_model_dimensions(self):
+        return self.model_dimensions
+
+    def forward(
+        self,
+        hidden_states: torch.Tensor,
+        position_embeddings: Tuple[torch.Tensor, torch.Tensor],
+        attention_mask: Optional[torch.Tensor],
+        past_key_value: Optional[Cache] = None,
+        cache_position: Optional[torch.LongTensor] = None,
+        **kwargs: Unpack[FlashAttentionKwargs],
+    ) -> Tuple[torch.Tensor, Optional[torch.Tensor], Optional[Tuple[torch.Tensor]]]:
+        input_shape = hidden_states.shape[:-1]
+        hidden_shape = (*input_shape, -1, self.head_dim)
+
+        query_states = self.q_proj(hidden_states).view(hidden_shape).transpose(1, 2)
+        key_states = self.k_proj(hidden_states).view(hidden_shape).transpose(1, 2)
+        value_states = self.v_proj(hidden_states).view(hidden_shape).transpose(1, 2)
+
+        cos, sin = position_embeddings
+        query_states, key_states = apply_rotary_pos_emb(
+            query_states, key_states, cos, sin
+        )
+
[…]

### 2025-05-10 src/module/llama.py

@@ -1,0 +1,102 @@
+from transformers.models.llama.modeling_llama import (
+    LlamaAttention,
+    apply_rotary_pos_emb,
+    eager_attention_forward,
+    logger,
+    ALL_ATTENTION_FUNCTIONS,
+    repeat_kv,
+)
+import torch
+from typing import Tuple, Optional, Unpack, Callable, Dict, List, Set
+import numpy as np
+from transformers.cache_utils import Cache
+from transformers.modeling_flash_attention_utils import FlashAttentionKwargs
+
+
+class LlamaBlockedAttention(LlamaAttention):
+    def __init__(
+        self,
+        config,
+        layer_idx,
+        disable_idx: Optional[Set[int]] = {},
+        disable_positional_encoding: Optional[bool] = False,
+    ):
+        super().__init__(config, layer_idx)
+        self.disable_idx = set(disable_idx)
+        self.disable_positional_encoding = disable_positional_encoding
+        self.save_model_dimensions = True if layer_idx == 0 else False
+        self.model_dimensions = None
+
+    def disable_head(self, head_idx: int):
+        self.disable_idx.add(head_idx)
+
+    def enable_head(self, head_idx: int):
+        if head_idx in self.disable_idx:
+            self.disable_idx.remove(head_idx)
+
+    def get_model_dimensions(self):
+        return self.model_dimensions
+
+    def forward(
+        self,
+        hidden_states: torch.Tensor,
+        position_embeddings: Tuple[torch.Tensor, torch.Tensor],
+        attention_mask: Optional[torch.Tensor],
+        past_key_value: Optional[Cache] = None,
+        cache_position: Optional[torch.LongTensor] = None,
+        **kwargs: Unpack[FlashAttentionKwargs],
+    ) -> Tuple[torch.Tensor, Optional[torch.Tensor], Optional[Tuple[torch.Tensor]]]:
+        
+        input_shape = hidden_states.shape[:-1]
+        hidden_shape = (*input_shape, -1, self.head_dim)
+
+        query_states = self.q_proj(hidden_states).view(hidden_shape).transpose(1, 2)
+        key_states = self.k_proj(hidden_states).view(hidden_shape).transpose(1, 2)
+        value_states = self.v_proj(hidden_states).view(hidden_shape).transpose(1, 2)
+        cos, sin = position_embeddings
+
+        if not self.disable_positional_encoding:
+            query_states, key_states = apply_rotary_pos_emb(
+                query_states, key_states, cos, sin
[…]

### 2025-05-10 src/module/qwen.py

@@ -1,0 +1,96 @@
+from transformers.models.qwen3.modeling_qwen3 import (
+    Qwen3Attention,
+    Qwen3Config,
+    Qwen3RMSNorm,
+    eager_attention_forward,
+    apply_rotary_pos_emb,
+    ALL_ATTENTION_FUNCTIONS,
+)
+import torch
+import torch.nn as nn
+from typing import Tuple, Optional, Unpack, Callable, Dict, List, Set
+from transformers.cache_utils import Cache
+from transformers.modeling_flash_attention_utils import FlashAttentionKwargs
+from transformers.utils import logging
+import numpy as np
+
+logger = logging.get_logger(__name__)
+
+
+class QwenBlockedAttention(Qwen3Attention):
+    """Multi-headed attention from 'Attention Is All You Need' paper"""
+
+    def __init__(self, config: Qwen3Config, layer_idx: int, disable_idx: Set[int] = {}, disable_positional_encoding: bool = False):
+        super().__init__(config, layer_idx)
+        self.disable_idx = set(disable_idx)
+        self.disable_positional_encoding = disable_positional_encoding
+        self.save_model_dimensions = True if layer_idx == 0 else False
+        self.model_dimensions = None
+
+    def disable_head(self, head_idx: int):
+        self.disable_idx.add(head_idx)
+
+    def enable_head(self, head_idx: int):
+        self.disable_idx.remove(head_idx)
+
+    def get_model_dimensions(self):
+        return self.model_dimensions
+
+    def forward(
+        self,
+        hidden_states: torch.Tensor,
+        position_embeddings: Tuple[torch.Tensor, torch.Tensor],
+        attention_mask: Optional[torch.Tensor],
+        past_key_value: Optional[Cache] = None,
+        cache_position: Optional[torch.LongTensor] = None,
+        **kwargs: Unpack[FlashAttentionKwargs],
+    ) -> Tuple[torch.Tensor, Optional[torch.Tensor], Optional[Tuple[torch.Tensor]]]:
+        input_shape = hidden_states.shape[:-1]
+        hidden_shape = (*input_shape, -1, self.head_dim)
+
+        query_states = self.q_norm(self.q_proj(hidden_states).view(hidden_shape)).transpose(1, 2)
+        key_states = self.k_norm(self.k_proj(hidden_states).view(hidden_shape)).transpose(1, 2)
+        value_states = self.v_proj(hidden_states).view(hidden_shape).transpose(1, 2)
+
[…]

### 2025-05-10 src/t.py

@@ -1,5 +1,6 @@
 from module.blocked_attention import load_altered_attention_model
-from util.data import load_numerical_data, load_grammar_data_by_category, load_word_problems_data
+from util.data import load_numerical_data, load_grammar_data_by_category
+from util.numerical import load_word_problems_data
 import torch
 from random import shuffle
 from module.algo import layer_select, head_select
@@ -7,73 +8,57 @@ import json
 from pprint import pprint
 
 if __name__ == "__main__":
-    numerical_data = load_numerical_data(margin=50)
-    shuffle(numerical_data)
-    numerical_data = numerical_data[:50]
+    
 
-    noun_agreement_data = load_grammar_data_by_category(
-        category="linguistics_term", value="determiner_noun_agreement", sample_rate=1
-    )
-    shuffle(noun_agreement_data)
-    noun_agreement_data = noun_agreement_data[:50]
+    device = torch.device("cuda:5" if torch.cuda.is_available() else "cpu")
 
-    word_problems_data = load_word_problems_data()
-
-    device = torch.device("cuda:6" if torch.cuda.is_available() else "cpu")
+    # Load the model
+    model_name = "qwen/qwen3-8b"
+    model_type = "qwen"
     model, tokenizer = load_altered_attention_model(
-        "google/gemma-2-9b-it", device, model_type="gemma"
-    )
-    with open("outputs/gemma_removed_layers_numerical.json", "r") as f:
-        numerical_layer_data = json.load(f)
-
-    with open("outputs/gemma_removed_layers_word_problems.json", "r") as f:
-        word_problems_layer_data = json.load(f)
-
-    with open("outputs/gemma_removed_layers_noun_agreement.json", "r") as f:
-        noun_agreement_layer_data = json.load(f)
-
-    head_select(
-        word_problems_data,
-        model,
-        tokenizer,
-        device,
-        layer_data=word_problems_layer_data,
-        metric_type="raw",
-        problem_type="word_problems",
-        use_system_message=False,
-        outfile="outputs/gemma_removed_head_pool_word_problems.json",
+        model_name, device, model_type=model_type
     )
 
-    # layer_select(
-    #     numerical_data,
-    #     model,
-    #     tokenizer,
-    #     device,
-    #     metric_type="raw",
-    #     problem_type="numerical",
-    #     use_system_message=False,
-    #     percentages=[1/20],
-    #     outfile="outputs/gemma_removed_layers_numerical.json",
-    # )
-    # layer_select(
-    #     word_problems_data,
-    #     model,
[…]

### 2025-05-10 src/util/data.py

@@ -4,6 +4,7 @@ import os
 import numpy as np
 from tqdm import tqdm
 from random import uniform
+from util.numerical import load_word_problems_data
 
 def load_grammar_data(file_path):
     with open(file_path, "r") as f:
@@ -78,18 +79,21 @@ def load_numerical_data(
     return ret_data
 
 
-def load_word_problems_data(
-    data_path: str = "../data/word_problems/fill_in_addition_subtraction_simple.jsonl",
-    use_tqdm: bool = False,
-):
-    data = []
-    with open(data_path, "r") as f:
-        for line in f:
-            try:
-                data.append(json.loads(line))
-            except:
-                print(line)
-    return data
+# def load_word_problems_data(
+#     data_path: str = "../data/word_problems/fill_in_addition_subtraction_simple.jsonl",
+#     use_tqdm: bool = False,
+# ):
+#     data = []
+#     with open(data_path, "r") as f:
+#         for line in f:
+#             try:
+#                 data.append(json.loads(line))
+#             except:
+#                 print(line)
+#     return data
+
+# def load_word_problems_data():
+#     return load_word_problems_data()
 
 
 def load_hidden_data(model_name, grammar_name, data_path="../data/blimp-hidden"):

### 2025-05-10 src/util/numerical.py

@@ -54,7 +54,7 @@ def fill_in_example(example, a, b, ans):
     return ex
 
 
-def load_word_problems_data(data_path: str = "../../data/word_problems/fill_in_simple.json", numerical_data_path: str = "../../data/arithmetic/addition_subtraction.jsonl", margin: int = 50):
+def load_word_problems_data(data_path: str = "../data/word_problems/template.json", numerical_data_path: str = "../data/arithmetic/addition_subtraction.jsonl", margin: int = 50):
     with open(data_path, "r") as f:
         data = json.load(f)
     with open(numerical_data_path, "r") as f:
@@ -92,9 +92,9 @@ def load_word_problems_data(data_path: str = "../../data/word_problems/fill_in_s
 
     shuffle(dataset)
 
-    with open('../../data/word_problems/fill_in_addition_subtraction_simple.jsonl', 'w') as f:
-        for item in dataset:
-            f.write(json.dumps(item) + '\n')
+    # with open('../../data/word_problems/fill_in_addition_subtraction_simple.jsonl', 'w') as f:
+    #     for item in dataset:
+    #         f.write(json.dumps(item) + '\n')
 
     return dataset
 

### 2025-05-10 src/module/algo.py
backup

@@ -94,14 +94,21 @@ def layer_select(
     return results
 
 
-def save_head_pool(head_pool, outfile):
+def save_head_pool(head_pool, outfile, starting_acc, lowest_acc, i, base_acc):
     head_dict = {}
     for (l, h) in head_pool:
         if l not in head_dict:
             head_dict[l] = []
         head_dict[l].append(h)
+    res_dict = {
+        "base_acc": base_acc,
+        "starting_acc": starting_acc,
+        "lowest_acc": lowest_acc,
+        "i": i,
+        "head_pool": head_dict,
+    }
     with open(outfile, "w") as f:
-        json.dump(head_dict, f, indent=4)
+        json.dump(res_dict, f, indent=4)
 
 def get_total_disabled_heads(model):
     disabled_heads = get_disabled_heads(model)
@@ -133,9 +140,9 @@ def head_select(
     least_impact = max(layer_impact)
     impacts = [(least_impact - v) for v in layer_impact]
     upper_quartile = np.percentile(impacts, percentile)
-    layer_pool = [i for i, v in enumerate(layer_impact) if (least_impact - v) > upper_quartile]
+    layer_pool = [i for i, v in enumerate(layer_impact) if (least_impact - v) >= upper_quartile]
     head_pool = [(l, h) for l in layer_pool for h in range(n_heads)]
-    K = int(2 * np.ceil(np.sqrt(len(head_pool))))
+    K = int(len(head_pool) / 2)
 
     base_acc = metric(data, model, tokenizer, device, metric_type, problem_type, D=D)
     print('Base Acc: ', base_acc)
@@ -176,7 +183,7 @@ def head_select(
             if acc < lowest_acc:
                 lowest_acc = acc
                 lowest_acc_head_batch = head_batch
-
+        
         print('Lowest Acc: ', lowest_acc, '| Head Batch: ', lowest_acc_head_batch)
     
         for l in range(n_layers):
@@ -188,14 +195,20 @@ def head_select(
                 break
             K = int(np.ceil(K / 2))
             print('Reducing K to ', K)
-            continue
 
-        save_head_pool(head_pool, outfile)
+        else:
+            save_head_pool(head_pool, outfile, starting_acc, lowest_acc, i, base_acc)
+            new_head_pool = []
+            for l, h in head_pool:
+                if (l, h) not in lowest_acc_head_batch:
+                    new_head_pool.append((l, h))
 
-        new_head_pool = []
-        for l, h in head_pool:
[…]

### 2025-05-10 src/module/qwen.py

@@ -31,7 +31,8 @@ class QwenBlockedAttention(Qwen3Attention):
         self.disable_idx.add(head_idx)
 
     def enable_head(self, head_idx: int):
-        self.disable_idx.remove(head_idx)
+        if head_idx in self.disable_idx:
+            self.disable_idx.remove(head_idx)
 
     def get_model_dimensions(self):
         return self.model_dimensions

### 2025-05-10 src/t.py

@@ -6,15 +6,21 @@ from random import shuffle
 from module.algo import layer_select, head_select
 import json
 from pprint import pprint
+import os
+
 
 if __name__ == "__main__":
     
 
-    device = torch.device("cuda:5" if torch.cuda.is_available() else "cpu")
+    device = torch.device("cuda:7" if torch.cuda.is_available() else "cpu")
 
     # Load the model
-    model_name = "qwen/qwen3-8b"
-    model_type = "qwen"
+    # model_name = "qwen/qwen3-8b"
+    # model_type = "qwen"
+    # model_type = "gemma"
+    # model_name = "google/gemma-2-9b-it"
+    model_type = "llama"
+    model_name = "meta-llama/Llama-3.1-8B-Instruct"
     model, tokenizer = load_altered_attention_model(
         model_name, device, model_type=model_type
     )
@@ -24,11 +30,11 @@ if __name__ == "__main__":
         shuffle(numerical_data)
         numerical_data = numerical_data[:50]
 
-        noun_agreement_data = load_grammar_data_by_category(
-            category="linguistics_term", value="determiner_noun_agreement", sample_rate=1
+        quantifiers_data = load_grammar_data_by_category(
+            category="linguistics_term", value="quantifiers", sample_rate=1
         )
-        shuffle(noun_agreement_data)
-        noun_agreement_data = noun_agreement_data[:50]
+        shuffle(quantifiers_data)
+        quantifiers_data = quantifiers_data[:50]
 
         word_problems_data = load_word_problems_data()
         shuffle(word_problems_data)
@@ -36,21 +42,27 @@ if __name__ == "__main__":
 
         data = {
             "numerical": numerical_data,
-            "noun_agreement": noun_agreement_data,
+            "quantifiers": quantifiers_data,
             "word_problems": word_problems_data,
         }
-
+    
         for problem_type, data in data.items():
-            layer_data = layer_select(
-                data,
-                model,
-                tokenizer,
-                device,
-                metric_type="raw",
-                problem_type=problem_type,
-                k=5,
-                outfile=f"layers/{model_type}/{problem_type}_{i}.json",
-            )
+            layer_data_file = f"layers/{model_type}/{problem_type}_{i}.json"
+            if not os.path.exists(layer_data_file):
+                layer_data = layer_select(
+                    data,
+                    model,
+                    tokenizer,
+                    device,
+                    metric_type="raw",
+                    problem_type=problem_type,
+                    k=5,
[…]

### 2025-05-10 src/util/chat.py

@@ -435,10 +435,16 @@ def load_embeddings_dataset(
                     }
                 )
             if "raw" in types_:
+                if problem_type == 'grammar':
+                    chat = "Pay attention to the grammatical correctness of the following sentence: {}"
+                elif problem_type == 'arithmetic':
+                    chat = "Pay attention to the arithmetic correctness of the following expression: {}"
+                elif problem_type == 'word_problems':
+                    chat = "Pay attention to the numerical correctness of the following statement: {}"
                 embeddings["raw"].append(
                     {
                         "states": get_eos_states(
-                            model, tokenizer, device, example["good"]
+                            model, tokenizer, device, chat.format(example["good"])
                         ),
                         "answer": 1,
                     }
@@ -446,7 +452,7 @@ def load_embeddings_dataset(
                 embeddings["raw"].append(
                     {
                         "states": get_eos_states(
-                            model, tokenizer, device, example["bad"]
+                            model, tokenizer, device, chat.format(example["bad"])
                         ),
                         "answer": 0,
                     }

### 2025-05-13 old2/chat1.py
backup

@@ -1,0 +1,631 @@
+from warnings import filterwarnings
+import logging
+from util.data import load_grammar_data
+from random import sample
+
+from tqdm import tqdm
+from random import shuffle
+from transformers import pipeline
+import torch
+import numpy as np
+from pprint import pprint
+from util.model import clear_cuda_cache
+
+logging.getLogger("transformers").setLevel(logging.ERROR)
+filterwarnings("ignore")
+
+np.random.seed(42)
+
+"""
+===== GRAMMAR =====
+"""
+
+__system_message_multiple_choice_grammar = """You are a helpful assistant that will help me figure out which sentence is grammatically correct and native-like.
+You should only output the number of the correct sentence, '1' if the first sentence is correct, '2' if the second sentence is correct.
+
+ONLY OUTPUT THE NUMBER OF THE CORRECT SENTENCE, NOTHING ELSE.
+
+For example:
+Which sentence is more grammatical and native-like?
+1) The cat is on the mat.
+2) The mat on is the cat.
+
+2"""
+
+__system_message_binary_grammar = """You are a helpful assistant that will help me figure out if a sentence is grammatically correct and native-like. 
+You should only output '1' if the sentence is grammatically correct and native-like, and '0' if it is not.
+
+ONLY OUTPUT '1' OR '0', NOTHING ELSE.
+
+For example:
+Is this sentence grammatically correct?
+1) The cat is on the mat.
+
+1"""
+
+__question_format_multiple_choice_grammar = """Which sentence is grammatically correct? 
+1) {sentence_1} 
+2) {sentence_2}"""
+
+__question_format_binary_grammar = """Is this sentence grammatically correct? 
+1) {sentence_1}"""
+
+"""
+===== ARITHMETIC =====
+"""
+
+__system_message_multiple_choice_arithmetic = """You are a helpful assistant that will help me determine which of two arithmetic expressions is correct.
+
+ONLY OUTPUT THE NUMBER OF THE CORRECT EXPRESSION (1 OR 2), NOTHING ELSE.
+For example:
+Which of the following arithmetic expressions is correct?
+1) 2 + 2 = 4
+2) 2 + 2 = 5
+
+1"""
+
+__system_message_binary_arithmetic = """You are a helpful assistant that will help me determine if an arithmetic expression is correct.
+
+ONLY OUTPUT '1' OR '0', NOTHING ELSE.
+For example:
+Is this arithmetic expression correct?
[…]

### 2025-05-13 src/module/algo.py

@@ -19,6 +19,7 @@ import json
 from util.data import load_numerical_data
 from random import shuffle
 from pprint import pprint
+from util.model import clear_cuda_cache
 
 
 def metric(
@@ -28,7 +29,7 @@ def metric(
     device,
     metric_type: str,
     problem_type: str,
-    D: int = 10,
+    D: int = 3,
 ):
     embeddings = load_embeddings_dataset(
         data,
@@ -91,10 +92,11 @@ def layer_select(
         
         with open(outfile, "w") as f:
             json.dump(results, f, indent=4)
+        clear_cuda_cache()
     return results
 
 
-def save_head_pool(head_pool, outfile, starting_acc, lowest_acc, i, base_acc):
+def save_head_pool(head_pool, outfile, starting_acc, lowest_acc, i, base_acc, terminated: bool = False, k_history: list = None):
     head_dict = {}
     for (l, h) in head_pool:
         if l not in head_dict:
@@ -106,6 +108,8 @@ def save_head_pool(head_pool, outfile, starting_acc, lowest_acc, i, base_acc):
         "lowest_acc": lowest_acc,
         "i": i,
         "head_pool": head_dict,
+        "terminated": terminated,
+        "k_history": k_history
     }
     with open(outfile, "w") as f:
         json.dump(res_dict, f, indent=4)
@@ -125,10 +129,11 @@ def head_select(
     device, 
     epsilon=.1,
     layer_data=None,
-    n_splits=10,
+    n_splits=20,
     metric_type="raw",
     problem_type="numerical",
-    D: int = 10,
+    D: int = 3,
+    T: int = 20,
     percentile: int = 75,
     outfile="outputs/gemma_numerical_head_pool.json",
 ):
@@ -152,25 +157,29 @@ def head_select(
             disable_head(model, l, h)
     
     starting_acc = metric(data, model, tokenizer, device, metric_type, problem_type, D=D)
+    delta = base_acc - starting_acc
     print('Starting Acc: ', starting_acc)
 
-    for l in layer_pool:
+    for l in range(n_layers):
         for h in range(n_heads):
             enable_head(model, l, h)
     
+    k_history = []
     i = 0
     while True:
+        clear_cuda_cache()
         pool_len = len(head_pool)
         if pool_len < K:
-            if K == 1:
+            if K == T:
                 break
-            K = int(np.ceil(K / 2))
+            K = max(T, int(np.ceil(K / 2)))
             print('Reducing K to ', K)
             continue
 
[…]

### 2025-05-13 src/module/blocked_attention.py

@@ -46,6 +46,7 @@ def load_altered_attention_model(
                 [] if idx not in disable_idx else disable_idx[idx],
             )
         blocked_attn.load_state_dict(layer.self_attn.state_dict())
+        del layer.self_attn
         layer.self_attn = blocked_attn
         clear_cuda_cache()
     model.to(device)

### 2025-05-13 src/t.py

@@ -4,74 +4,110 @@ from util.numerical import load_word_problems_data
 import torch
 from random import shuffle
 from module.algo import layer_select, head_select
+from util.model import clear_cuda_cache
 import json
 from pprint import pprint
 import os
-
+import sys
+import uuid
 
 if __name__ == "__main__":
     
+    idx = 0
+    device = torch.device(f"cuda:{idx}" if torch.cuda.is_available() else "cpu")
 
-    device = torch.device("cuda:7" if torch.cuda.is_available() else "cpu")
+    while True:
+        free_memory, total_memory = torch.cuda.mem_get_info(device)
+        if free_memory / total_memory < 0.9:
+            idx += 1
+            device = torch.device(f"cuda:{idx}" if torch.cuda.is_available() else "cpu")
+        else:
+            break
 
-    # Load the model
-    # model_name = "qwen/qwen3-8b"
-    # model_type = "qwen"
-    # model_type = "gemma"
-    # model_name = "google/gemma-2-9b-it"
-    model_type = "llama"
-    model_name = "meta-llama/Llama-3.1-8B-Instruct"
+    print(f"Using device: {device}")
+
+    model_type = sys.argv[1]
+    problem_type = sys.argv[2]
+
+    
+    if model_type == "gemma":
+        model_name = "google/gemma-2-9b-it"
+    elif model_type == "llama":
+        model_name = "meta-llama/Llama-3.1-8B-Instruct"
+    else:
+        raise ValueError(f"Invalid model type: {model_type}")
     model, tokenizer = load_altered_attention_model(
         model_name, device, model_type=model_type
     )
 
-    for i in range(10):
+    iterations = 10
+    batch_size = 50
+    print(f"Running {model_type} model")
+    for i in range(iterations):
         numerical_data = load_numerical_data(margin=50)
         shuffle(numerical_data)
-        numerical_data = numerical_data[:50]
+        numerical_data = numerical_data[:batch_size]
 
-        quantifiers_data = load_grammar_data_by_category(
-            category="linguistics_term", value="quantifiers", sample_rate=1
+        determiner_noun_agreement_data = load_grammar_data_by_category(
+            category="linguistics_term", value="determiner_noun_agreement", sample_rate=1
         )
-        shuffle(quantifiers_data)
-        quantifiers_data = quantifiers_data[:50]
+        shuffle(determiner_noun_agreement_data)
+        determiner_noun_agreement_data = determiner_noun_agreement_data[:batch_size]
 
[…]

### 2025-05-13 src/util/chat.py

@@ -16,6 +16,10 @@ filterwarnings("ignore")
 
 np.random.seed(42)
 
+"""
+===== GRAMMAR =====
+"""
+
 __system_message_multiple_choice_grammar = """You are a helpful assistant that will help me figure out which sentence is grammatically correct and native-like.
 You should only output the number of the correct sentence, '1' if the first sentence is correct, '2' if the second sentence is correct.
 
@@ -39,12 +43,16 @@ Is this sentence grammatically correct?
 
 1"""
 
-__question_format_multiple_choice_grammar = """Which sentence is more grammatically correct? 
+__question_format_multiple_choice_grammar = """Which sentence is grammatically correct? 
 1) {sentence_1} 
 2) {sentence_2}"""
 
 __question_format_binary_grammar = """Is this sentence grammatically correct? 
 1) {sentence_1}"""
+
+"""
+===== ARITHMETIC =====
+"""
 
 __system_message_multiple_choice_arithmetic = """You are a helpful assistant that will help me determine which of two arithmetic expressions is correct.
 
@@ -56,10 +64,6 @@ Which of the following arithmetic expressions is correct?
 
 1"""
 
-__question_format_multiple_choice_arithmetic = """Which of the following arithmetic expressions is correct?
-1) {sentence_1}
-2) {sentence_2}"""
-
 __system_message_binary_arithmetic = """You are a helpful assistant that will help me determine if an arithmetic expression is correct.
 
 ONLY OUTPUT '1' OR '0', NOTHING ELSE.
@@ -69,8 +73,16 @@ Is this arithmetic expression correct?
 
 1"""
 
-__question_format_binary_arithmetic = """Is this arithmetic expression correct?
+__question_format_multiple_choice_arithmetic = """Which of the following arithmetic expressions is numerically correct?
+1) {sentence_1}
+2) {sentence_2}"""
+
+__question_format_binary_arithmetic = """Is this arithmetic expression numerically correct?
 1) {sentence_1}"""
+
+"""
+===== WORD PROBLEMS =====
+"""
 
 __system_message_multiple_choice_word_problems = """You are a helpful assistant that will help me figure out which statement is correct.
 You should only output the number of the correct statement, '1' if the first statement is correct, '2' if the second statement is correct.
@@ -98,7 +110,7 @@ Is this statement correct?
 __question_format_binary_word_problems = """Is this statement numerically correct?
 1) {sentence_1}"""
 
[…]
